\subsection{Embedding axioms}
\begin{definition}
    A large cardinal axiom \( \Phi\mathsf{C} \) is called an \emph{embedding axiom} if \( \Phi(\kappa) \) holds if and only if there is a transitive model \( M \) and elementary embedding \( j : \mathrm{V}_\lambda \to M \) with critical point \( \kappa \) with certain additional properties.
\end{definition}
\( \mathsf{M}(\kappa) \) is the simplest embedding axiom.
The remaining large cardinal axioms in this course will take the form of embedding axioms.
\begin{definition}
    An embedding \( j : \mathrm{V}_\lambda \to M \) with critical point \( \kappa \) is called \emph{\( \beta \)-strong} if \( \mathrm{V}_{\kappa + \beta} \subseteq M \).
    A cardinal \( \kappa \) is called \emph{\( \beta \)-strong} if there is a \( \beta \)-strong embedding with critical point \( \kappa \).
\end{definition}
\( \beta \)-stable properties are preserved by \( \beta \)-strong embeddings.
In particular, by the reflection argument, if \( \Phi \) is \( \beta \)-stable and \( \kappa \) is \( \beta \)-strong with \( \Phi(\kappa) \), then \( \kappa \) is the \( \kappa \)th cardinal with property \( \Phi \).

Note that \( \kappa \) is measurable if and only if \( \kappa \) is 1-strong, and if \( \kappa \) is 2-strong then \( \qty{\alpha < \kappa \mid \mathsf{M}(\alpha)} \) and \( \qty{\alpha < \kappa \mid \mathsf{Surv}(\alpha)} \) are of size \( \kappa \).
If we write \( \beta\mathsf{-S}(\kappa) \) to denote that \( \kappa \) is \( \beta \)-strong, then
\[ \mathsf{SurvC} <_{\Con} 2\mathsf{-S}(\kappa) \]
This also gives an example of \( j_{U_j} \neq j \), as the ultrapower embedding of any ultrafilter is never 2-strong.
\begin{definition}
    A large cardinal property \( \Phi \) is said to have \emph{witness objects} of rank \( \beta \) if there is a formula \( \Psi \) that is downwards absolute for transitive models such that
    \[ \Phi(\kappa) \leftrightarrow \forall x.\, \exists y \in \mathrm{V}_{\kappa + \beta}.\, \Psi(x, y, \kappa) \]
\end{definition}
Any large cardinal property with witness objects of rank \( \beta \) is \( \beta \)-stable.
\begin{example}
    \begin{enumerate}
        \item Weakly compact cardinals have witness objects of rank 1: for all colourings, there exists a homogeneous set in \( \mathrm{V}_{\kappa + 1} \).
        \item Measurable cardinals have witness objects of rank 2: there is a \( \kappa \)-complete nonprincipal ultrafilter on \( \kappa \).
        The initial \( \forall x \) quantifier is not needed in this case.
        \item Surviving cardinals also have witness objects of rank 2, namely, a pair of ultrafilters.
    \end{enumerate}
\end{example}
In particular, inaccessibility is 0-stable, weak compactness is 1-stable, and measurability and survivability are 2-stable.

Observe that if \( \beta \)-strong cardinals have witness objects, they cannot be of rank \( \beta \).

\subsection{Removing the inaccessible}
The ultrapower constructions used an inaccessible cardinal above a measurable cardinal, so that we could obtain a set-sized universe containing a measurable cardinal.
When trying to do this with the real universe, we encounter several problems.
\begin{enumerate}
    \item The definition of ultrapowers requires a set model.
    \item In the fundamental theorem of measurable cardinals, we have a quantification over \( j \) and \( M \).
    If these are proper classes, this quantification cannot be expressed in the usual language of set theory.
    \item Also, in the fundamental theorem of measurable cardinals, we use the notion of an elementary embedding, which is only definable for set models.
\end{enumerate}
To solve problem (i), we would like to construct \( \faktor{\mathrm{V}^\kappa}{\sim_U} \).
Note that \( \mathrm{V}^\kappa \) is a well-defined class; it is the class of all functions with domain \( \kappa \).
For such functions, it is easy to define the equivalence relation \( \sim_U \).
However, the equivalence classes \( [f]_U = \qty{g \in \mathrm{V}^\kappa \mid f \sim_U g} \) are all proper classes.
So \( \faktor{\mathrm{V}^\kappa}{\sim_U} \) is no longer a standard class; classes containing proper classes are typically not allowed.
This can be resolved using \emph{Scott's trick}.
If \( C \) is a nonempty class, then there is a minimal \( \alpha \) such that \( C \cap \mathrm{V}_\alpha \neq \varnothing \).
This is a nonempty set.
Define \( \operatorname{scott}(C) = C \cap \mathrm{V}_\alpha \) for this \( \alpha \).
Hence, if \( [f]_U \neq [g]_U \), we have \( \operatorname{scott}([f]_U) \neq \operatorname{scott}([g]_U) \).
We can therefore define
\[ \faktor{\mathrm{V}^\kappa}{U} = \qty{\operatorname{scott}([f]_U) \mid \dom f = \kappa} \]
To obtain our model \( M \), we took the Mostowski collapse of \( \faktor{{\mathrm{V}_\lambda}^\kappa}{U} \).
Therefore, we need a class version of the Mostowski collapse.
Recall that a relation \( E \subseteq C \times C \) is \emph{set-like} if for all \( x \in C \), the class \( \qty{y \in C \mid y \mathrel{E} x} \) is a set.
\begin{theorem}
    Let \( C \) be a class, and let \( E \subseteq C \times C \) be a binary relation on \( C \) that is well-founded, extensional, and set-like.
    Then there is a unique transitive class \( T \) such that \( (T, \in) \cong (C, E) \).
\end{theorem}
This may be proven in an almost identical fashion to Mostowski's collapsing theorem for sets.

For problems (ii) and (iii), recall that the fundamental theorem of measurable cardinals was that \( \mathsf{M}(\kappa) \) is equivalent to the statement that there is an elementary embedding \( j : \mathrm{V}_\lambda \to M \) with critical point \( \kappa \).
Measurability is witnessed at \( \kappa + 2 \), but the elementary embedding is not witnessed anywhere below \( \lambda \), so we cannot extend this definition to the usual universe.
We can solve this by extending our set theory to an appropriate class theory.
Standard class theories include \emph{von Neumann--Bernays--G\"odel} or \( \mathsf{NBG} \), and \emph{Morse--Kelley} or \( \mathsf{MK} \).
These theories have very different notions of class.
\( \mathsf{NBG} \) set theory is based upon the idea that definable formulas give the classes.
It is a `minimal class theory' where all classes are definable.
\( \mathsf{MK} \) is based on the idea that \( \mathrm{Ord} \) behaves externally like in inaccessible cardinal.
In this theory, there could be undefinable classes, and more classes than sets.

This resolves problem (ii), as we are permitted to work in a language in which we may quantify over proper classes.
However, this does not solve problem (iii).
Elementarity cannot be expressed as a single formula, but becomes a schema.
This causes additional problems as we need the existential over \( j \) and \( M \) to be part of each formula.
This could be solved by extending the language to add symbols for \( j \) and \( M \).
Another resolution is to observe that \( \Sigma_1 \)-elementarity suffices, as is explored in Kanamori's book \emph{The Higher Infinite} on page 45.
This can be defined using a single formula, therefore solving problem (iii).
