\subsection{Products of primes}
\begin{proposition}
	Every \(n \geq 2\) is expressible as a product of primes.
\end{proposition}
\begin{proof}
	We use induction on an integer \(n\), starting at 2, a trivial case.
	Given \(n > 2\), we have two cases:
	\begin{itemize}
		\item \(n\) is prime.
		      Therefore, \(n\) is a product of primes as required.
		\item \(n\) is composite.
		      We know that \(n\) can be split into two factors, denoted here as \(a\), \(b\).
		      Using (strong) induction, we know that because both \(a\) and \(b\) are smaller than \(n\), they are expressible as a product of primes.
		      We simply multiply these products together to express \(n\) as a product of primes.
	\end{itemize}
\end{proof}

\subsection{Infinite primes}
\begin{proposition}
	There are infinitely many primes.
\end{proposition}
\begin{proof}
	Assume there exists a largest prime.
	Then, the list of primes is \(p_1, p_2 \cdots p_k\).
	Let \(n=p_1 p_2 \cdots p_k + 1\).
	Then \(n\) has no prime factor.
	This is a contradiction immediately because we know that every number greater than two has a factorisation, but this doesn't.
\end{proof}

We want to prove that prime factorisation is unique (up to the ordering).
We need that \(p \mid ab \implies p \mid a \lor p \mid b\).
However, this is hard to answer --- \(p\) is defined in terms of what divides it, not what it divides.
This is the reverse of its definition, so we need to prove it in a more round-about way.

\subsection{Highest common factors}
For \(a, b \in \mathbb N\), a number \(c \in \mathbb N\) is defined to be the highest common factor if:
\begin{itemize}
	\item \(c \mid a\) and \(c \mid b\), and
	\item For all other factors \(d\) (\(d \mid a\) and \(d \mid b\)), we have that \(d \mid c\).
\end{itemize}
The second point implies that it is the \textit{highest} common factor, but it is actually slightly stronger.
Note that, for example, if a pair's common factors were 1, 2, 3, 4, 6 then the numbers would not have a highest common factor, because 4 does not divide 6.

\subsection{Division algorithm}
The Division Algorithm allows us to write any number \(n \in \mathbb N\) as a multiple \(q\in\mathbb N\) of \(k\in \mathbb N\) with some remainder \(r\in\mathbb N\) such that \(0 \leq r < k\); this can be shortened to \(n = qk + r\).
We begin by writing 1 in this form: \(1 = 0k + 1\).
Inductively, \(n\) can be written as:
\[
	n = (n-1) + 1 = q_0 k + r_0 + 1
\]
where \(q_0\) and \(r_0\) are the results of \(q\) and \(r\) for \(n-1\).
Note that we have two cases:
\begin{itemize}
	\item If \(r_0 + 1 < k\): the result is simply \(n = q_0k + (r_0+1)\)
	\item Else (\(r_0 + 1 = k\)): the result is \(n = (q_0 + 1)k + 0\)
\end{itemize}

\subsection{Euclid's algorithm}
We can find the highest common factor of two natural numbers \(a\) and \(b\) (without loss of generality, we assume that \(a \leq b\)).
\begin{itemize}
	\item Write \(a\) as some multiple \(q_1\) of \(b\), with remainder \(r_1\).
	\item Write \(b\) as some multiple \(q_2\) of \(r_1\), with remainder \(r_2\).
	\item Write \(r_1\) as some multiple \(q_3\) of \(r_2\), with remainder \(r_3\).
	\item Continue until \(r_{n+1}=0\).
	      Then, \(r_n\) is the highest common factor of \(a\) and \(b\).
	      We know that the algorithm terminates because \(r_k < r_{k-1}\) so it will terminate in at most \(b\) steps.
\end{itemize}
We now prove that the algorithm works.
\begin{proof}
	We need to prove that it is a common factor and then that it divides all other common factors.
	\begin{itemize}
		\item On the last line of the algorithm, we have \(r_{n-1} = q_{n+1} r_n + 0\), so we know that \(r_n \mid r_{n-1}\).
		      On the second last line, we have \(r_{n-2} = q_n r_{n-1} + r_n\), but \(r_n\) divides \(r_{n-1}\), so \(r_n\) must divide \(r_{n-2}\).
		      We can continue this logic up to the start of the algorithm, where we can see that \(r_n \mid a\) and \(r_n \mid b\).
		      So \(r_n\) is a common factor of \(a\) and \(b\).
		\item Given some other common factor \(d \neq r_n\), we can look at the first line of the algorithm to see that \(d \mid r_1\).
		      Using this, we can use the next line to see that \(d \mid r_2\).
		      Continuing to the last line, we have \(d \mid r_n\).
	\end{itemize}
	So \(r_n\) is the highest common factor of \(a\) and \(b\).
	Therefore, the highest common factor exists and is unique for any natural numbers \(a\) and \(b\).
\end{proof}

\subsection{Linear combinations}
Consider running Euclid's algorithm on the numbers 87 and 52.
\begin{align*}
	87 & = 1 \cdot 52 + 35 \\
	52 & = 1 \cdot 35 + 17 \\
	35 & = 2 \cdot 17 + 1  \\
	17 & = 17 \cdot 1 + 0
\end{align*}
1 is the highest common factor of 87 and 52.
Now, we can write 1 as a linear combination of 87 and 52 by looking at each line of this algorithm in the reverse direction (ignoring the bottom line).
\begin{align*}
	1 & = 35 - 2 \cdot 17                         \\
	  & = 35 - 2 \cdot (52 - 1 \cdot 35)          \\
	  & = -2 \cdot 52 + 3 \cdot 35                \\
	  & = -2 \cdot 52 + 3 \cdot (87 - 1 \cdot 52) \\
	  & = 3 \cdot 87 - 5 \cdot 52
\end{align*}
Each two lines of this equation represents one line on Euclid's algorithm.
We end up with a linear combination of the two input numbers.
We can prove that this linear combination exists in the general case.
\begin{theorem}
	Let \(a, b \in \mathbb N\).
	Then there exist some \(x, y \in \mathbb Z\) such that \(xa + yb = \HCF(a, b)\).
\end{theorem}
\begin{proof}
	Run Euclid's algorithm on \(a\) and \(b\), and let the output be \(r_n\).
	Then we have \(r_n = x r_{n-1} + y r_{n-2}\) for some \(x, y \in \mathbb Z\).
	So, \(r_n\) can be written as a linear combination of \(r_{n-1}\) and \(r_{n-2}\).
	Also, from the previous line we know that \(r_{n-1} = x r_{n-2} + y r_{n-3}\) for some other \(x\) and \(y\).
	So we can rewrite \(r_{n}\) as a linear combination of \(r_{n-2}\) and \(r_{n-3}\).
	Inductively, we can rewrite \(r_n\) as a linear combination of \(a\) and \(b\) by moving up the lines of the algorithm.
\end{proof}
We can also make an alternate proof without using Euclid's algorithm.
Note that this algorithm does not show how to generate this linear combination, it just shows that one exists.
\begin{proof}[Alternate Proof]
	Let \(h\) be the least positive linear combination of \(a\) and \(b\).
	We want to prove that \(h = \HCF (a, b)\).
	\begin{itemize}
		\item Assume that there exists some common factor \(d\) of \(a\) and \(b\), so that \(d\mid a\) and \(d\mid b\).
		      Then for some \(x\) and \(y\), \(d \mid (xa + yb)\).
		      So \(d \mid h\).
		\item Suppose \(h\) does not divide \(a\).
		      Then \(a = qh + r\) where \(q\) is the quotient and \(r\) is the remainder (\(r \neq 0\)).
		      Then \(r = a - qh = a - q(xa + yb)\) for some integers \(x\) and \(y\).
		      So \(r\) is a linear combination of \(a\) and \(b\).
		      But this is a contradiction because we said that \(h\) was the smallest one.
		      So \(h\) divides \(a\).
	\end{itemize}
	Therefore \(h\) is the highest common factor.
\end{proof}

\subsection{Linear diophantine equations}
Suppose \(a\), \(b\) and \(c\) are natural numbers.
When can we solve \(ax + by = c\) for \(x, y \in \mathbb Z\)?
Well, by looking at the previous theorem, we might guess that \(c\) must be some multiple of the highest common factor of \(a\) and \(b\).
This can be proven in the general case.
\begin{corollary}[Bezout's Theorem]
	Let \(a, b, c \in \mathbb N\).
	Then \(ax + by = c\) where \(x, y \in \mathbb Z\) has a solution if and only if \(\HCF(a, b) \mid c\).
\end{corollary}
\begin{proof}
	Let \(h = \HCF(a, b)\).
	We must prove this bi-implication in both directions.
	\begin{itemize}
		\item First, let us assume that \(ax+by=c\) has a solution for some integers \(x\) and \(y\).
		      Since \(h \mid a\) and \(h \mid b\) then \(h \mid (ax+by)\) so \(h \mid c\).
		\item Conversely, we know that \(h = ax + by\) for some \(x\) and \(y\) by the above theorem.
		      We can multiply both sides by the integer \(c/h\) (this is an integer because \(h \mid c\)).
		      Then we have an expression for \(c\) as a linear combination of \(a\) and \(b\) as required.
	\end{itemize}
\end{proof}

\subsection{Fundamental theorem of arithmetic}
\begin{lemma}
	Let \(p\) be a prime, let \(a, b \in \mathbb N\).
	Then \(p \mid ab\) implies \(p \mid a\) or \(p \mid b\).
\end{lemma}
\begin{proof}
	Let \(p \mid ab\).
	Then we have two cases, either \(p\) divides \(a\) or it does not divide \(a\).
	If it does, our statement is trivially true.
	Otherwise, we want to prove that \(p\) divides \(b\).

	Now \(\HCF(p, a)=1\) as \(p\) is a prime, and it does not divide \(a\).
	So 1 can be written as some linear combination of \(p\) and \(a\): \(px + ay = 1\) for some \(x, y \in \mathbb Z\).

	Now we can multiply both sides by \(b\), giving \(pbx + aby = b\).
	Since \(p\) divides \(ab\), \(p\) must divide the left hand side.
	So \(p\) divides \(b\).
\end{proof}
Note that we started with a kind of `negative' statement: `\(p\) does not divide \(a\)'; this told us that we cannot do something (namely, factorise it).
We turned it into a `positive' statement: `\(px + ay = 1\)'; this allows us to rearrange to find out information about these variables.
Converting `negative' statements to `positive' statements is a useful tool in making proofs.

\begin{theorem}[Fundamental Theorem of Arithmetic]
	Every \(n \in \mathbb N\) is uniquely expressible as a product of primes.
\end{theorem}
\begin{proof}
	Note that we have already proven that a prime factorisation is possible in Section 3.4; we just need to prove uniqueness of a factorisation (at least, down to its order).
	We will use induction on some integer \(n\) that we wish to factorise.
	Clearly the theorem is true for \(n=1\) (assuming empty products are valid) and \(n=2\).

	So given that \(n > 2\) we suppose that there exist two possible factorisations:
	\[
		n = p_1 p_2 \cdots p_k = q_1 q_2 \cdots q_l
	\]
	We want to prove that \(k=l\) and that (after reordering) \(p_i = q_i\) for all valid \(i\).

	We know that \(p_1 \mid n\), so \(p_1 \mid (q_1 \cdots q_l)\).
	So there must exist some \(i\) where \(p_1 \mid q_i\).
	But since \(q_i\) is prime, \(p_1 = q_i\).
	Let us reorder the list such that \(q_i\) is moved to the front, so that \(p_1 = q_1\).
	\[
		n = p_1 p_2 \cdots p_k = p_1 q_2 \cdots q_l
	\]
	Now, we divide the entire equation by \(p_1\) to give
	\[
		\frac{n}{p_1} = p_2 \cdots p_k = q_2 \cdots q_l
	\]
	The integer \(\frac{n}{p_1}\) is smaller than \(n\), so we can use induction to assume that its factorisation is unique.
	Therefore
	\[
		[p_2, p_3 \cdots p_k] = [q_2, q_3 \cdots q_l]
	\]
	So the prime factorisation of \(n\) is unique.
\end{proof}

\subsection{Highest common factor}
The common factors of two numbers \(m = p_1^{a_1} \cdots p_k^{a_k}\) and \(n = p_1^{b_1} \cdots p_k^{b_k}\) where \(a\) and \(b\) are zero or above is given by \(p_1^{c_1} \cdots p_k^{c_k}\) where \(c_i \leq \min(a_i, b_i)\) So the highest common factor is given by \(c_i = \min(a_i, b_i)\).

\subsection{Lowest common multiple}
The common multiples of those two numbers is given by \(d_i \geq \max(a_i, b_i)\).
So analogously the lowest common multiple is given by \(d_i = \max(a_i, b_i)\).

We have an interesting property that \(\HCF(m, n) \LCM(m, n) = mn\).
This is true because any term \(p_i\) is given by \(p_i^{\min(a_i, b_i)}p_i^{\max(a_i, b_i)} = p_i^{a_i + b_i}\).
