\subsection{Introduction}
We have a notion of `size' for finite sets.
Is there such an analogous notion for infinite sets?
We will say that a set \(X\) is countable if \(X\) is finite, or it bijects with \(\mathbb N\).
Equivalently, we can list out the elements of the set, and each element will appear in the list.
Here are some examples.
\begin{enumerate}
	\item Clearly any finite set is countable.
	\item \(\mathbb N\) is countable.
	\item \(\mathbb Z\) is countable, let us construct the list of numbers
	      \[
		      0, 1, -1, 2, -2, 3, -3, 4, -4, \dots
	      \]
\end{enumerate}
It makes sense now to consider two sets to have the same size if they biject with each other.

\subsection{Countability under injections}
\begin{proposition}
	A set \(X\) is countable if and only if it injects into \(\mathbb N\).
\end{proposition}
\begin{proof}
	The forward implication is trivial: if \(X\) is finite, then there must be an injection in to \(\mathbb N\), and if it bijects with \(\mathbb N\) then that bijection is a valid injection.
	This encompasses both cases of countable sets.

	Now let us consider the reverse implication.
	We may assume \(X\) is infinite, since if \(X\) is finite then by definition \(X\) is countable.
	We know that \(X\) injects onto \(\mathbb N\) under some injective function \(f\), so \(X\) bijects with \(\Im f\).
	So it is enough to show that the image \(\Im f\) is countable.
	We will now set \(a_1\) to be the least element of \(\Im f\), and \(a_2\) to be the least element not equal to \(a_1\), and so on.
	In general, \(a_n = \min (\Im f \setminus \{ a_i : 0 \leq i < n \})\).
	Then \(\Im f\) is the set \(\{ a_1, a_2, \dots \}\).
	There are no extra elements that we have not covered, since each \(a \in X\) is \(a_n\) for some \(n\), because \(a=a_n, n \leq a\).
	So we have listed elements of \(\Im f\), so \(\Im f\) is countable, so \(X\) is countable.
\end{proof}
Thus, we can view countability as being `at most as large as \(\mathbb N\)'.
For instance, any subset of a countable set is also countable.

\begin{remark}
	In \(\mathbb R\), let
	\[
		X = \left\{ \frac{1}{2}, \frac{2}{3}, \frac{3}{4}, \dots \right\} \cup \{ 1 \}
	\]
	Then \(X\) is countable, as we can list it as
	\[
		1, \frac{1}{2}, \frac{2}{3}, \frac{3}{4}, \dots
	\]
	But if we counted from `least element' to `most element', we would never reach the element 1 in countable time.
	Note further that if we find it difficult to construct a list for a set, it does not mean it is uncountable, it could just mean that we haven't found the right list yet.
\end{remark}

\subsection{Products of countably infinite sets}
\begin{theorem}
	\(\mathbb N \times \mathbb N\) is countable.
\end{theorem}
\begin{proof}[Proof 1]
	We will define \(a_1 = (1, 1)\), and inductively define
	\[
		a_n = \begin{cases}
			(p-1, q+1) & \text{if } p > 1 \\
			(q+1, 1)   & \text{if } p = 1
		\end{cases}
	\]
	where \(a_{n-1} = (p, q)\).
	Therefore, we are essentially moving across antidiagonals of the plane.
	This does hit every point \((x, y) \in \mathbb N \times \mathbb N\), for example by induction on \(x+y\), so we have listed all elements of \(\mathbb N \times \mathbb N\).
\end{proof}
\begin{proof}[Proof 2]
	If we can define an injective function \(\mathbb N \times \mathbb N \to \mathbb N\), then it is countable.
	For example, let \(f = 2^x 3^y\).
	\(f\) is injective, so \(\mathbb N \times \mathbb N\) is countable.
\end{proof}

\subsection{Countable unions of countable sets}
Proof 2 is also a way to show the following theorem:
\begin{theorem}
	Let \(A_1, A_2, A_3, \dots\) be countable sets.
	Then \(A_1 \cup A_2 \cup A_3 \cup \dots\) is countable.
	Less formally, `a countable union of countable sets is countable'.
\end{theorem}
\begin{proof}
	For each \(i\), \(A_i\) is countable, so we can list \(A_i\) as \(a_{i1}, a_{i2}, a_{i3}, \dots\) which may or may not terminate.
	We can then define
	\[
		f\colon \bigcup_{n \in \mathbb N}A_n \to \mathbb N;\quad f(x) = 2^i 3^j
	\]
	where \(x = a_{ij}\).
	If \(x\) is in more than one set, just take the least \(i\) that is valid.
	Then \(f\) is an injection so the union is countable.
\end{proof}

\subsection{Partitioning into countable subsets}
Here are some examples of using this theorem by partitioning sets as a countable union of countable subsets.
\begin{enumerate}
	\item \(\mathbb Q\) is countable, since it is a countable union of countable sets:
	      \[
		      \mathbb Q = \mathbb Z \cup \frac{1}{2}\mathbb Z \cup \frac{1}{3}\mathbb Z \cup \dots
	      \]
	      Each \(\frac{1}{n}\mathbb Z\) is countable, since they biject with \(\mathbb Z\) which is a countable set.
	      It doesn't matter if we've counted an element in \(\mathbb Q\) twice; the above theorem works even with intersecting sets.
	\item The set \(\mathbb A\) of all algebraic numbers is countable.
	      It is enough to show that the set of integer polynomials is countable, since each polynomial has a finite amount of roots and then \(\mathbb A\) is a countable union of finite sets.
	      Now, to show that the set of integer polynomials is countable, it is enough to show that for each degree \(d\) it is countable, since it is a countable union of all polynomials of degree \(d\) (again using the above theorem).
	      To specify a polynomial of degree \(d\) you must name its coefficients, so this set injects into \(\mathbb Z^{d+1}\), so we must just show that \(\mathbb Z^{d+1}\) is countable (not a bijection since the first term of the polynomial must be nonzero).
	      We know that \(\mathbb Z^n\) is countable because we can inductively show that \(\mathbb Z^2, \mathbb Z^3, \mathbb Z^4, \dots\) are countable inductively.
\end{enumerate}

\subsection{Uncountable sets}
\begin{definition}
	A set is uncountable if there is no way to count the set.
\end{definition}
\begin{theorem}
	\(\mathbb R\) is uncountable.
\end{theorem}
\begin{proof}[Proof (Cantor's Diagonal Argument)]
	We will show that \((0, 1)\) is uncountable, then clearly \(\mathbb R\) is uncountable.
	Suppose \((0, 1)\) is countable.
	Then given a sequence \(r_1, r_2, \dots\) in \((0, 1)\), we just need to find some number \(s \in (0, 1)\) not contained within this sequence.
	For each \(r_n\), we have a decimal expansion \(r_n = 0.r_{n1}r_{n2}r_{n3}\dots\).
	Let us now write all of these numbers in a matrix-style form:
	\begin{align*}
		r_1 & = 0.r_{11}r_{12}r_{13}\dots \\
		r_2 & = 0.r_{21}r_{22}r_{23}\dots \\
		r_3 & = 0.r_{31}r_{32}r_{33}\dots \\
		\vdots
	\end{align*}
	We just need to construct some number \(s\) that is not in this list.
	So, let us simply make sure that for any given \(r\) value, there is at least one digit that does not match.
	The easiest way to construct such a number is
	\[
		s = 0.s_1 s_2 s_3 \dots
	\]
	where \(s_1 \neq r_{11}\), \(s_2 \neq r_{22}\), \(s_3 \neq r_{33}\) and so on.
	We can pick any numbers we like according to these constraints, but we should avoid picking digits 0 and 9 since \(0.1000\dots = 0.0999\dots\) for example, which could cause unnecessary ambiguity.
	Then \(s \neq r_1, s \neq r_2, \dots\) since there is at least one mismatched digit in the expansion for each \(r_i\); they differ in decimal digit \(i\).
	So \(\mathbb R\) is uncountable.
\end{proof}
This is another proof that transcendental numbers exist.
\(\mathbb R\) is uncountable and \(\mathbb A\) is countable, so there exists a transcendental number.
Indeed, `most' numbers are transcendental, i.e.\ \(\mathbb R \setminus \mathbb A\) is uncountable (because if \(\mathbb R \setminus \mathbb A\) were countable, then \(\mathbb R\) would be \((\mathbb R \setminus \mathbb A) \cup \mathbb A\) which is a finite union of countable sets \contradiction).

\subsection{Countability of power sets}
\begin{theorem}
	The power set \(\mathcal P(\mathbb N)\) is uncountable.
\end{theorem}
\begin{proof}
	Suppose the subsets of \(\mathbb N\) are listed as \(S_1, S_2, S_3, \dots\) then we want to construct another set \(S\) that is not equal to any of the other sets \(S_i\).
	So for each set \(S_i\), we must ensure that \(S\) and \(S_i\) differ for at least one value.
	An easy way to do this is to include the number \(i\) in the subset if \(S_i\) does not contain the number, and to exclude \(i\) if \(i \in S_i\).
	Then \(S\) differs from \(S_i\) at position \(i\).
	This is the same logic as the diagonal argument above.
	We have:
	\[
		S = \{ n \in \mathbb N : n \notin S_n \}
	\]
	So \(S\) is not on the list \(S_1, S_2, S_3, \dots\) no matter what way we choose to list the elements, so \(\mathcal P(\mathbb N)\) is uncountable.
\end{proof}
\begin{remark}
	Alternatively, we could just inject \((0, 1)\) into \(\mathcal P(\mathbb N)\).
	For example, consider \(x \in (0, 1)\) represented as \(0.x_1x_2x_3x_4\dots\) in binary where the \(x_1, x_2, \dots\) are zero or one (not ending with an infinite amount of 1s).
	We can convert this \(x\) into a subset of \(\mathbb N\) by considering the set \(\{ n \in \mathbb N : x_n = 1 \}\).
	Then the uncountability follows.
\end{remark}
In fact, our proof of this theorem shows the following.
\begin{theorem}
	For any set \(X\), there is no bijection from \(X\) to the power set \(\mathcal P(X)\).
\end{theorem}
For example, \(\mathbb R\) does not biject with \(\mathcal P(\mathbb R)\).
The proof in fact will show that there is no surjection from \(X\) to its power set; i.e.\ the power set is `larger' than \(X\).
\begin{proof}
	Given any function \(f\colon X \to \mathcal P(X)\), we will show \(f\) is not surjective.
	Let \(S = \{ x \in X: x \notin f(x) \}\).
	Then \(S\) does not belong to the image of \(f\) because they differ at element \(x\); for all \(x\) we have \(S \neq f(x)\).
\end{proof}
\begin{remark}
	Note that:
	\begin{enumerate}
		\item This is similar in some sense to Russell's paradox.
		\item This theorem gives another proof that there is no universal set \(\mathscr E\), since its power set \(\mathcal P(\mathscr E) \subseteq \mathscr E\).
		      But of course, there is always a surjection from a set to a subset.
		      This is a contradiction.
	\end{enumerate}
\end{remark}

\subsection{Disjoint real intervals}
This is an example on countability.
Let \(A_i, i \in I\) be a family of open, pairwise disjoint intervals.
Must this family be countable?
Note that it is not as simple as just listing from left to right, for example consider
\[
	\left(\frac{1}{2}, 1\right), \left(\frac{1}{3}, \frac{1}{2}\right), \left(\frac{1}{4}, \frac{1}{3}\right), \dots, (-1, 0)
\]
Then the leftmost interval is \((-1, 0)\), but there is no `next interval' just after it.
Also consider
\[
	\left( 0, \frac{1}{2} \right), \left( \frac{1}{2}, \frac{2}{3} \right), \left( \frac{2}{3}, \frac{3}{4} \right), \dots, (1, 2)
\]
Then we can list the first infinitely many intervals, but we will never reach \((1, 2)\).
The answer turns out to be true; the family is countable.
Here are two important proofs.
\begin{proof}[Proof 1]
	Each interval \(A_i\) contains a rational number \(a_i\).
	The rationals \(\mathbb Q\) are countable.
	So let us just list the \(a_i\).
	The family is therefore countable.
\end{proof}
\begin{proof}[Proof 2]
	\(\{ i \in I: A_i \text{ has length } \leq 1\}\) is certainly countable, since it injects into \(\mathbb Z\) (here, as all \(A_i\) contain some integer).
	Further, \(\left\{ i \in I: A_i \text{ has length } \leq \frac{1}{2} \right\}\) is countable for the same reason.
	Essentially, for all \(n\), \(\left\{ i \in I: A_i \text{ has length } \leq \frac{1}{n} \right\}\) is countable.
	We have written all the intervals as a countable union (over \(n\)) of countable sets.
\end{proof}

\subsection{Summary}
To show a set \(X\) is uncountable:
\begin{enumerate}
	\item Run a diagonal argument; or
	\item Inject an uncountable set into \(X\)
\end{enumerate}
To show a set \(X\) is countable:
\begin{enumerate}
	\item List all the elements (usually fiddly); or
	\item Inject \(X\) into \(\mathbb N\) (or another countable set); or
	\item Express \(X\) as a countable union of countable sets (usually the best); or
	\item If \(X\) is `in' or `near' \(\mathbb R\), consider \(\mathbb Q\) (see Proof 2 above).
\end{enumerate}
