\subsection{Sets}
A set is any* collection of mathematical objects.
\((\forall x, x \in A \iff x \in B) \iff (A = B)\).
In words, two sets which have the same members are considered to be the same; order of members is not important in a set.
There is no `multiple membership' of a set, \(\{ a, a \} = \{ a \}\).

\subsection{Subsets}
Given a set \(A\) and a property \(p(x)\), we can form \(\{ x \in A: p(x) \}\); the subset of all members of \(A\) with property \(p\).
This is sometimes called the `subset selection' rule or axiom.
We can say that \(B\) is a subset of \(A\) if \(\forall x, x \in B \implies x \in A\), written \(B \subseteq A\).
Further, \(A = B \iff A \subseteq B, B \subseteq A\).

\subsection{Unions and Intersections}
Given sets \(A\) and \(B\), we can form their union \(A \cup B = \{ x: x \in A \lor x \in B \}\).
We can also form their intersection \(A \cap B = \{ x: x \in A \wedge x \in B \}\).
If \(A \cap B = \varnothing\), we say \(A\) and \(B\) are disjoint.
Note that we could consider \(A \cap B\) as a special case of subset selection; the subset of \(A\) with the property that the element is in \(B\).
Therefore, \(A \cap B \subseteq A\), and \(A \cap B \subseteq B\).
We define the set difference \(A \setminus B = \{ x \in A: x \notin B \}\).

Note that \(\cap\) and \(\cup\) are commutative and associative.
Also, \(\cup\) is distributive over \(\cap\), and \(\cap\) is distributive over \(\cup\).
For example, let us prove that \(A \cap (B \cup C) = (A \cap B) \cup (A \cap C)\).
\begin{itemize}
	\item (LHS \(\subseteq\) RHS) Given \(x \in A \cap (B \cup C)\), we have \(x \in A\) and also either \(x \in B\) or \(x \in C\).
	      If \(x \in B\) then \(x \in A \cap B\) so \(x \in (A \cap B) \cup (A \cap C)\); and vice versa for \(C\).
	\item (RHS \(\subseteq\) LHS) Given \(x \in (A \cap B) \cup (A \cap C)\), either \(x \in A \cap B\) or \(x \in A \cap C\).
	      If \(x \in A \cap B\) then \(x \in A\) and \(x \in B \cup C\) as required; and vice versa for the other case.
\end{itemize}
As the union is associative, we can have bigger unions of more sets.
For example, if we let \(A_n = \{ n^2, n^3 \}\) for each \(n \in \mathbb N\), the infinite union
\[
	A_1 \cup A_2 \cup A_3 \cup \cdots = \bigcup_{n=1}^\infty A_n = \bigcup_{n \in \mathbb N} A_n = \{ x \in N: x \text{ is a square or a cube} \}
\]
When we use the \(n \in \mathbb N\) on the large union symbol, we call \(\mathbb N\) the `index set'.
Note that the infinite union is not defined as a limit of finite unions; it is simply defined using set comprehension.
In general, given a set \(I\), and sets \(A_i\), \(i \in I\), we can form
\[
	\bigcup_{i \in I}A_i = \{ x: \exists i \in I, x \in A_i \}
\]
and
\[
	\bigcap_{i \in I}A_i = \{ x: \forall i \in I, x \in A_i \}
\]
Note that we cannot form an intersection when \(I = \varnothing\), as will be explained later.
