\subsection{Ordered Pairs}
For any \(a, b\), we can form the ordered pair \((a, b)\), where equality is checked component-wise.
For sets \(A, B\), we can form their product \(A \times B = \{ (a, b) : a \in A, b \in B \}\).
For example, \(\mathbb R^2 = \mathbb R \times \mathbb R\) can be viewed as a plane.
We can form other sizes of tuples similarly.

\subsection{Power Sets}
For any set \(X\), we can form the power set \(\mathcal P(X)\) consisting of all subsets of \(X\).
\[
	\mathcal P(X) = \{ Y: Y \subseteq X \}
\]
For example:
\[
	\mathcal P(\{ 1, 2 \}) = \{ \varnothing, \{ 1 \}, \{ 2 \}, \{ 1, 2\} \}
\]

\subsection{Russell's Paradox}
For a set \(A\), we can always form the set \(\{ x \in A: p(x) \}\) for any property \(p\).
We cannot, however, form the set \(\{ x: p(x) \}\).
Suppose we could form such a set, then we could form the set \(X = \{ x: x \notin x \}\).
Now, is \(X \in X\)? If this is true, then it fails the defining property \(x \notin x\).
If this is false, then the defining property is true, so it must be in the set.
This is a contradiction in both cases.

Similarly, there is no `universal' set \(\mathscr E\), meaning \(\forall x, x \in \mathscr E\).
Otherwise we could form the \(X\) above by \(\{ x \in \mathscr E: p(x) \}\).
To guarantee that a given set exists, we need to obtain it in some way from known sets.

\subsection{Finite Sets}
We will write \(\mathbb N_0 = \mathbb N \cup \{ 0 \}\).
For \(n \in \mathbb N_0\), we can say that a set \(A\) has size \(n\) if we can write \(A = \{ a_1, a_2, \cdots, a_n \}\) where the \(a_i\) are distinct.
A set is called finite if it has a size \(n \in \mathbb N_0\).

Note that a set cannot have size \(n\) and size \(m\) for \(n \neq m\).
Suppose that \(A\) has size \(n\) and size \(m\) where \(n, m > 0\).
Then, removing an element, we obtain a set that has size \(n-1\) and \(m-1\).
By induction on the larger of \(n\) and \(m\), we will eventually reach a size of both zero and non-zero which is a contradiction.

\begin{proposition}
	A set of size \(n\) has exactly \(2^n\) subsets.
\end{proposition}
\begin{proof}[Proof 1]
	We may assume that our set is simply \(\{ 1, 2, \cdots, n \}\) by relabelling.
	When constructing a subset \(S\) from this set, there are \(n\) independent binary choices for whether a given element should be within this subset, since for example either \(1 \in S\) or \(1 \notin S\) must be true.
	So there are \(2^n\) distinct choices of subset you can make.
\end{proof}
\begin{proof}[Proof 2]
	We will prove this inductively on \(n\), noting that \(n=0\) is trivial.
	For any subset \(T \subseteq \{ 1, 2, \cdots n-1 \}\), how many \(S \subseteq \{ 1, \cdots, n \}\) have \(S \cap \{ 1, 2, \cdots n-1 \} = T\)? Exactly two: \(T\) and \(T \cup \{ n \}\).
	So there are two choices for how to extend this subset to the new element \(n\).
	So the number of subsets is \(2 \cdot 2^{n-1} = 2^n\).
\end{proof}
\noindent In some sense Proof 2 is a more `formal' version of Proof 1, using induction rather than intuition.
We sometimes say that if \(A\) has size \(n\), then \(\abs{A} = n\), and that \(A\) is an \(n\)-set.

\subsection{Binomial Coefficients}
For \(n \in \mathbb N_0\) and \(0 \leq k \leq n\), we can write \(\binom{n}{k}\) for the number of subsets of an \(n\)-set that are of size \(k\).
\[
	\binom{n}{k} = \abs{\left\{ S \subseteq \{ 1, 2, \dots, n \}: \abs{S} = k \right\}}
\]
For example, there are six 2-sets in a 4-set.
There is a formula for this, but generally this definition is a lot easier to use.
Note that \(\binom{n}{0} = 1\), \(\binom{n}{n} = 1\), and \(\binom{n}{1}=n\) where \(n>0\).

Note that \(\binom{n}{0} + \binom{n}{1} + \dots + \binom{n}{n} = 2^n\) as each side counts the number of subsets in an \(n\)-set.
Also:
\begin{enumerate}
	\item \(\binom{n}{k} = \binom{n}{n-k}\) (\(\forall n \in N_0, 0 \leq k \leq n\)).
	      Indeed, specifying which \(k\) members to pick for a subset is equivalent to specifying which \(n-k\) members not to pick.
	\item \(\binom{n}{k} = \binom{n-1}{k-1} + \binom{n-1}{k}\) (\(\forall n \in \mathbb N, 0 < k < n\)).
	      Indeed, the number of \(k\)-subsets of \(\{ 1, 2, \dots, n \}\) without \(n\) is \(\binom{n-1}{k}\).
	      The number of \(k\)-subsets of \(\{ 1, 2, \dots, n \}\) that do contain \(n\) is \(\binom{n-1}{k-1}\) as we must pick the remaining \(k-1\) elements of this new subset.
	      So in total, \(\binom{n-1}{k-1} + \binom{n-1}{k}\) encapsulates both possibilities.
\end{enumerate}
This last point illustrates that Pascal's Triangle will give all the binomial coefficients since it perfectly encapsulates the relationship between a given element of the triangle with two elements from the previous row.
The exact proof follows from the other known properties of the binomial coefficients.
