\subsection{Injection, Surjection and Bijection}
\begin{definition}
	A function $f\colon A \to B$ is
	\begin{itemize}
		\item injective, if $\forall a, a' \in A$, we have $a \neq a' \implies f(a) \neq f(a')$, or equivalently, $f(a) = f(a') \implies a = a'$, or in words, `different points stay different' (e.g. example 6 above).
		\item surjective, if $\forall b \in B$, $\exists a \in A$ such that $f(a) = b$, or in words, `everything in $B$ is hit' (e.g. examples 6 and 8).
		\item bijective, if it is injective and surjective, or in words, `everything in $B$ is hit exactly once', or `$f$ pairs up elements of $A$ and elements of $B$' (e.g. example 6, or $f\colon \mathbb R \to \mathbb R$ given by $f(x) = x^3$).
	\end{itemize}
\end{definition}
\begin{definition}
	For a function $f\colon A \to B$, $A$ is the domain, $B$ is the range, and $\{ b \in B : \exists a \in A \st f(a) = b \}$ is the image.
\end{definition}
We must always provide the domain and range of a function; a function's properties depend on this. For example, is the function $f$ defined by $f(x) = f^2$ injective? If $f\colon \mathbb N \to \mathbb N$, then it is injective, but if $f\colon \mathbb Z \to \mathbb Z$, then it is not.

There are a number of properties that hold specifically for finite sets $A$, $B$:
\begin{enumerate}
	\item There is no surjection $A \to B$ if $\abs{B} > \abs{A}$.
	\item There is no injection $A \to B$ if $\abs{A} > \abs{B}$.
	\item For a function $f\colon A \to A$, $f$ injective $\iff$ $f$ surjective. Hence, if $f$ is either injective or surjective, it is bijective.
	\item There is no bijection from $A$ to any proper subset of $A$.
\end{enumerate}
As counterexamples for infinite sets:
\begin{enumerate}
	\item We define $f_0\colon \mathbb N \to \mathbb N$ by $f_0(x) = x+1$. Then, $f_0$ is injective but not surjective.
	\item We define $f_1\colon \mathbb N \to \mathbb N$ by $f_0(x) = x-1$, or 1 if $x=1$. Then, $f_0$ is surjective but not injective.
	\item We define $g\colon \mathbb N \to \mathbb N \setminus \{ 1 \}$ by $g(x) = x+1$. Then, $g$ is bijective between $\mathbb N$ and a proper subset of $\mathbb N$.
\end{enumerate}

\subsection{More Examples of Functions}
\begin{enumerate}
	\item For any set $X$ we have $1_X\colon X \to X$ defined by $1_X(x) = x$. This is known as the identity function on $X$.
	\item For any set $X$, and $A \subset X$, we have an indicator function (or characteristic function) $\chi_A\colon X \to \{ 0, 1 \}$ defined by
	      \[ \chi_A(x) = \begin{cases}
			      0 & x \notin A \\
			      1 & x \in A
		      \end{cases} \]
	\item A sequence of reals $x_1, x_2, \dots$ is a function $f\colon \mathbb N \to \mathbb R$ defined by $f(n) = x_n$.
	\item The operation $+$ on $\mathbb N$ is a function $\mathbb N^2 \to \mathbb N$.
	\item A set $X$ has size $n$ $\iff$ there is a bijection between $X$ and $\{ 1, 2, \dots, n \}$.
\end{enumerate}

\subsection{Composition of Functions}
Given $f\colon A \to B$ and $g\colon B \to C$, we define the composition $g\circ f \colon A \to C$, given by $(g\circ f)(a) = g(f(a))$. For example, if $f\colon \mathbb R \to \mathbb R$, $f(x) = 2x$, $g\colon \mathbb R \to \mathbb R$, $g(x) = x+1$, then $(f \circ g)(x) = 2(x+1)$, and $(g \circ f)(x) = 2x + 1$.

In general, the operation $\circ$ is not commutative, as we can see from this example. However, $\circ$ is associative. Given $f\colon A \to B$, $g\colon B \to C$, $h\colon C \to D$, we have $h \circ (g \circ f) = (h \circ g) \circ f$. Indeed, for any input $x \in A$,
\[ (h \circ (g \circ f))(x) = h((g \circ f)(x)) = h(g(f(x))) = (h \circ g)(f(x)) = ((h \circ g)\circ f)(x) \]
Thus $(h \circ (g \circ f))(x) = ((h \circ g)\circ f)(x)$ for every $x \in A$, so $h \circ (g \circ f) = (h \circ g)\circ f$.
