\subsection{Countability of Power Sets}
\begin{theorem}
	The power set $\mathcal P(\mathbb N)$ is uncountable.
\end{theorem}
\begin{proof}
	Suppose the subsets of $\mathbb N$ are listed as $S_1, S_2, S_3, \dots$ then we want to construct another set $S$ that is not equal to any of the other sets $S_i$. So for each set $S_i$, we must ensure that $S$ and $S_i$ differ for at least one value. An easy way to do this is to include the number $i$ in the subset if $S_i$ does not contain the number, and to exclude $i$ if $i \in S_i$. Then $S$ differs from $S_i$ at position $i$. This is the same logic as the diagonal argument above. We have:
	\[ S = \{ n \in \mathbb N : n \notin S_n \} \]
	So $S$ is not on the list $S_1, S_2, S_3, \dots$ no matter what way we choose to list the elements, so $\mathcal P(\mathbb N)$ is uncountable.
\end{proof}
\begin{remark}
	Alternatively, we could just inject $(0, 1)$ into $\mathcal P(\mathbb N)$. For example, consider $x \in (0, 1)$ represented as $0.x_1x_2x_3x_4\dots$ in binary where the $x_1, x_2, \dots$ are zero or one (not ending with an infinite amount of 1s). We can convert this $x$ into a subset of $\mathbb N$ by considering the set $\{ n \in \mathbb N : x_n = 1 \}$. Then the uncountability follows.
\end{remark}
In fact, our proof of this theorem shows the following.
\begin{theorem}
	For any set $X$, there is no bijection from $X$ to the power set $\mathcal P(X)$.
\end{theorem}
For example, $\mathbb R$ does not biject with $\mathcal P(\mathbb R)$. The proof in fact will show that there is no surjection from $X$ to its power set; i.e. the power set is `larger' than $X$.
\begin{proof}
	Given any function $f\colon X \to \mathcal P(X)$, we will show $f$ is not surjective. Let $S = \{ x \in X: x \notin f(x) \}$. Then $S$ does not belong to the image of $f$ because they differ at element $x$; for all $x$ we have $S \neq f(x)$.
\end{proof}
\begin{remark}
	Note that:
	\begin{enumerate}
		\item This is similar in some sense to Russell's paradox.
		\item This theorem gives another proof that there is no universal set $\mathscr E$, since its power set $\mathcal P(\mathscr E) \subseteq \mathscr E$. But of course, there is always a surjection from a set to a subset. This is a contradiction.
	\end{enumerate}
\end{remark}

\subsection{Disjoint Real Intervals}
This is an example on countability. Let $A_i, i \in I$ be a family of open, pairwise disjoint intervals. Must this family be countable? Note that it is not as simple as just listing from left to right, for example consider
\[ \left(\frac{1}{2}, 1\right), \left(\frac{1}{3}, \frac{1}{2}\right), \left(\frac{1}{4}, \frac{1}{3}\right), \dots, (-1, 0) \]
Then the leftmost interval is $(-1, 0)$, but there is no `next interval' just after it. Also consider
\[ \left( 0, \frac{1}{2} \right), \left( \frac{1}{2}, \frac{2}{3} \right), \left( \frac{2}{3}, \frac{3}{4} \right), \dots, (1, 2) \]
Then we can list the first infinitely many intervals, but we will never reach $(1, 2)$. The answer turns out to be true; the family is countable. Here are two important proofs.
\begin{proof}[Proof 1]
	Each interval $A_i$ contains a rational number $a_i$. The rationals $\mathbb Q$ are countable. So let us just list the $a_i$. The family is therefore countable.
\end{proof}
\begin{proof}[Proof 2]
	$\{ i \in I: A_i \text{ has length } \leq 1\}$ is certainly countable, since it injects into $\mathbb Z$ (here, as all $A_i$ contain some integer). Further, $\left\{ i \in I: A_i \text{ has length } \leq \frac{1}{2} \right\}$ is countable for the same reason. Essentially, for all $n$, $\left\{ i \in I: A_i \text{ has length } \leq \frac{1}{n} \right\}$ is countable. We have written all the intervals as a countable union (over $n$) of countable sets.
\end{proof}

\subsection{Summary of Countability}
To show a set $X$ is uncountable:
\begin{enumerate}
	\item Run a diagonal argument; or
	\item Inject an uncountable set into $X$
\end{enumerate}
To show a set $X$ is countable:
\begin{enumerate}
	\item List all the elements (usually fiddly); or
	\item Inject $X$ into $\mathbb N$ (or another countable set); or
	\item Express $X$ as a countable union of countable sets (usually the best); or
	\item If $X$ is `in' or `near' $\mathbb R$, consider $\mathbb Q$ (see Proof 2 above).
\end{enumerate}
