\subsection{Proper Time}
A particle moves along a trajectory \(\vb x(t)\). The velocity of this particle is \(\dv{\vb x}{t} = \vb u(t)\). The path in spacetime is parametrised by \(t\). Both \(\vb x\) and \(t\) vary under a Lorentz transformation. Now, consider a particle at rest in \(S'\) with \(\vb x' = 0\). The invariant interval on the world line is
\[ \Delta s^2 = c^2 \Delta t^2 \]
We define the proper time \(\tau\) as
\[ \Delta \tau = \frac{1}{c}\Delta s \]
In particular, in \(S'\), \(\Delta\tau = \Delta t\), so the proper time is the time experienced in the rest frame of the particle. However, the equation \(\Delta \tau = \frac{1}{c}\Delta s\) holds in all frames, since \(\Delta s\) is Lorentz invariant. Note further that \(\Delta s\) is real since this always represents a timelike interval, as it represents a particle travelling through spacetime. We can therefore instead parametrise this particle's world line by its proper time, rather than by considering the time in any particular frame. So \(\vb x\) and \(t\) are both functions of \(\tau\) in any given reference frame. Further, infinitesimal changes are related by
\begin{align*}
	\dd{\tau}               & = \frac{\dd{s}}{c}                                       \\
	                        & = \frac{1}{c}\sqrt{c^2\dd{t}^2 - \abs{\dd{\vb x}}^2}     \\
	                        & = \frac{1}{c}\sqrt{c^2\dd{t}^2 - \abs{\vb u}^2 \dd{t}^2} \\
	                        & = \qty(1 - \frac{\vb u^2}{c^2})^{\frac{1}{2}}\dd{t}      \\
	\therefore \dv{t}{\tau} & = \gamma_{\vb u}
\end{align*}
where \(\gamma_{\vb u} = \qty(1 - \frac{\vb u^2}{c^2})^{\frac{1}{2}}\). Now, the total time observed by a particle moving along its world line is
\[ T = \int \dd{\tau} = \int \frac{\dd{t}}{\gamma_{\vb u}} \]

\subsection{4-Velocity}
We can parametrise the position 4-vector of a particle using \(\tau\), written
\[ X(\tau) = \begin{pmatrix}
		ct(\tau) \\ \vb x(\tau)
	\end{pmatrix} \]
We define the 4-velocity as
\[ U = \dv{\tau}X = \begin{pmatrix}
		c\dv*{t}{\tau} \\ \dv*{\vb x}{\tau}
	\end{pmatrix} = \dv{t}{\tau} \begin{pmatrix}
		c \\ \vb u
	\end{pmatrix} = \gamma_{\vb u} \begin{pmatrix}
		c \\ \vb u
	\end{pmatrix} \]
Since \(X' = \Lambda X\), we also have that
\[ U' = \Lambda U \]
because \(\tau\) is invariant. Note that any quantity whose components transform according to this rule is called a 4-vector, and in particular, the derivative of a 4-vector with respect to an invariant is also a 4-vector. Also, the scalar product \(U \cdot U\) is invariant under Lorentz transforms. Indeed, in the rest frame of a particle moving with 4-velocity \(U\), in this frame we have \(U = c^2\). In other frames,
\[ U \cdot U = \gamma^2 (c^2 - \vb u^2) = c^2 \]
as expected.

\subsection{Transformation of Velocities}
We have found that in special relativity, we cannot simply add velocities together. Consider a transformation \(\Lambda\) from \(S\) to \(S'\), where \(S'\) is moving (relative to \(S\)) at a speed \(v\) in the \(x\) direction. Consider a particle moving in \(S\) at speed \(u\) at an angle \(\theta\) to the \(x\) axis (with no component in the \(z\) axis). In \(S'\), it moves with speed \(u'\) at an angle \(\theta'\). We can write the 4-velocities as
\[ U = \begin{pmatrix}
		\gamma_{\vb u}c           \\
		\gamma_{\vb u}u\cos\theta \\
		\gamma_{\vb u}u\sin\theta \\
		0
	\end{pmatrix};\quad U' = \begin{pmatrix}
		\gamma_{\vb u'}c             \\
		\gamma_{\vb u'}u'\cos\theta' \\
		\gamma_{\vb u'}u'\sin\theta' \\
		0
	\end{pmatrix} \]
and further,
\[ U' = \Lambda U \]
where
\[ \Lambda = \begin{pmatrix}
		\gamma_v              & -\gamma_v \frac{v}{c} & 0 & 0 \\
		-\gamma_v \frac{v}{c} & \gamma_v              & 0 & 0 \\
		0                     & 0                     & 1 & 0 \\
		0                     & 0                     & 0 & 1
	\end{pmatrix} \]
Carrying out the matrix multiplication, we find
\[ \begin{pmatrix}
		\gamma_{\vb u'}c             \\
		\gamma_{\vb u'}u'\cos\theta' \\
		\gamma_{\vb u'}u'\sin\theta' \\
		0
	\end{pmatrix} = \begin{pmatrix}
		\gamma_v              & -\gamma_v \frac{v}{c} & 0 & 0 \\
		-\gamma_v \frac{v}{c} & \gamma_v              & 0 & 0 \\
		0                     & 0                     & 1 & 0 \\
		0                     & 0                     & 0 & 1
	\end{pmatrix} \begin{pmatrix}
		\gamma_{\vb u}c           \\
		\gamma_{\vb u}u\cos\theta \\
		\gamma_{\vb u}u\sin\theta \\
		0
	\end{pmatrix} \implies \left\{ \begin{array}{l}
		\displaystyle
		u'\cos\theta' = \frac{u\cos\theta - v}{1 - uv\cos\theta/c^2}      \\
		\displaystyle
		\tan\theta' = \frac{u\sin\theta}{\gamma_{\vb u}(u\cos\theta - v)} \\
	\end{array} \right. \]
The first equation corresponds to the normal transformation law for Lorentz transforms. The second equation, corresponding to a change in angle due to the motion of the observer, is called aberration. In particular, when \(u = c\), we can see that light rays appear to change direction due to the relative motion of the emitter and the observer.
