\subsection{Energy-Momentum 4-Vector}
We define the 4-momentum of a particle of mass $m$ and 4-velocity $U$ to be
\[ P = mU = m\gamma_{\vb u} \begin{pmatrix}
		c \\ \vb u
	\end{pmatrix} \]
Since $U$ is a 4-vector, we must have that $m$ is invariant under a Lorentz transformation. We will call this $m$ the `rest mass' of the object, defined as the mass as measured in the rest frame of the particle. The 4-momentum of a system of particles is defined as the sum of the 4-momenta of its individual particles. The spatial components of $P$, given by $\mu = 1, 2, 3$, can be referred to as the relativistic 3-momentum, given by $\vb p = \gamma_{\vb u} m \vb u$. This matches with the definition as seen in Newtonian physics, except that the mass $m$ is replaced by $\gamma_{\vb u} m$. We call this quantity the `apparent mass' of the particle or system of particles, as it represents the mass of the particle as observed by a different reference frame. Note that $\abs{\vb p}$ and $\gamma_{\vb u} m$ both tend to infinity as the particle approaches the speed of light. Note that the first component of $P$, $P^0$, is
\[ \gamma_{\vb u} mc = \frac{mc}{\sqrt{1 - \frac{\vb u^2}{c^2}}} = \frac{1}{c}\qty(mc^2 + \frac{1}{2}m\vb u^2 + \dots) \]
We recognise the $\frac{1}{2}m\vb u^2$ term as the kinetic energy of the particle. We interpret $P^0$ as an energy, divided by $c$ (to conserve units).
\[ P = \begin{pmatrix}
		\frac{1}{c} E \\ \vb p
	\end{pmatrix} \]
where
\[ E = \gamma_{\vb u} mc^2 = mc^2 + \frac{1}{2}m\vb u^2 + \dots \]
Note that as $\abs{\vb u} \to c$, $E \to \infty$. Since $P$ contains an energy term as well as a momentum term, we also call $P$ the energy-momentum 4-vector. Note that for a stationary particle of rest mass $m$, we have
\[ E = mc^2 \]
This implies that mass is a form of energy. The energy of a moving particle is
\[ E = mc^2 + \underbrace{(\gamma_{\vb u} - 1)mc^2}_{\mathclap{\text{relativistic kinetic energy}}} \]
Since $P \cdot P = \frac{E^2}{c^2} - \abs{\vb p}^2$ is Lorentz invariant, we have
\[ P \cdot P = m^2c^2 \]
Hence,
\[ \frac{E^2}{c^2} = \abs{\vb p}^2 + m^2c^2 \]
In Newtonian physics, mass is conserved, and energy is also conserved. In relativistic physics, mass is not conserved by itself, since it is a form of energy. From this derivation, it is theoretically possible to convert between mass and kinetic energy.

\subsection{Massless Particles}
A massless particle has zero rest mass. Such particles can have nonzero momentum and nonzero energy, because they are travelling at the speed of light, giving $\gamma_{\vb u} = \infty$. Since $P \cdot P = m^2c^2$, there are no factors of $\gamma$ in this expression giving
\[ P \cdot P = 0 \]
So such a particle travels along a light-like trajectory. Therefore there is no Lorentz transformation that brings a given reference frame into the rest frame of the particle, so we cannot define proper time for such a particle. Since $m^2c^2 = 0$, we must have
\[ \frac{E^2}{c^2} = \abs{\vb p}^2 \implies E = \abs{\vb p}c \]
Then,
\[ P = \frac{E}{c} \begin{pmatrix}
		1 \\ \vu n
	\end{pmatrix} \]
where $\vu n$ is a unit 3-vector in the direction of travel of the particle.

\subsection{Newton's Second Law}
Now that we have defined $P$ for all particles, we can rewrite Newton's second law in special relativity as
\[ \dv{P}{\tau} = F \]
where $F$ is the 4-force. If the 3-force is $\vb F$, we have
\[ F = \gamma_{\vb u} \begin{pmatrix}
		\vb F \cdot \vb u / c \\
		\vb F
	\end{pmatrix} \]
Hence,
\[ \dv{E}{\tau} = \gamma_{\vb u} \vb F \cdot \vb u;\quad \dv{\vb p}{\tau} = \gamma_{\vb u} \vb F \]
giving
\[ \dv{E}{t} = \vb F \cdot \vb u;\quad \dv{\vb p}{t} = \vb F \]
which are the familiar Newtonian expressions for rate of work and rate of change of momentum. We can now define 4-acceleration:
\[ F = mA \]
where $m$ is the rest mass. Hence,
\[ \dv{U}{\tau} = A \]

\subsection{Application of Special Relativity to Particle Physics}
In Newtonian physics, when two particles collide, we must consider the conservation of 3-momentum. In special relativity however, we must instead consider the conservation of 4-momentum:
\[ P = \begin{pmatrix}
		\frac{E}{c} \\ \vb p
	\end{pmatrix} \]
It is often convenient, when dealing with systems of particles, to let the origin of our frame of reference be the centre of momentum. This is the frame such that the total 3-momentum of the system is zero. However, this cannot be done when dealing with massless particles since there does not exist such a rest frame.

