\subsection{Probability Spaces and $\sigma$-algebras}
\begin{definition}
	Suppose $\Omega$ is a set, and $\mathcal F$ is a collection of subsets of $\Omega$. We call $\mathcal F$ a $\sigma$-algebra if
	\begin{enumerate}
		\item $\Omega \in \mathcal F$
		\item if $A \in \mathcal F$, then $\stcomp{A} \in \mathcal F$
		\item for any countable collection $(A_n)_{n \geq 1}$ with $A_n \in \mathcal F$ for all $n$, we must also have that $\bigcup_n A_n \in \mathcal F$
	\end{enumerate}
\end{definition}
\begin{definition}
	Suppose that $\mathcal F$ is a $\sigma$-algebra on $\Omega$. A function $\mathbb P \colon \mathcal F \to [0, 1]$ is called a probability measure if
	\begin{enumerate}
		\item $\prob{\Omega} = 1$
		\item for any countable disjoint collection of sets $(A_n)_{n \geq 1}$ in $\mathcal F$ ($A_n \in \mathcal F$ for all $n$), then $\prob{\bigcup_{n \geq 1}A_n} = \sum_{n \geq 1} \prob{A_n}$ (this is called `countable additivity')
	\end{enumerate}
	We say that $\prob{A}$ is `the probability of $A$'. We call $(\Omega, \mathcal F, \mathbb P)$ a probability space, where $\Omega$ is the sample space, $\mathcal F$ is the $\sigma$-algebra, and $\mathbb P$ is the probability measure.
\end{definition}

\begin{remark}
	When $\Omega$ is countable, we take $\mathcal F$ to be all subsets of $\Omega$, i.e. $\mathcal F = \mathcal P(\Omega)$, its power set.
\end{remark}
\begin{definition}
	The elements of $\Omega$ are called outcomes, and the elements of $\mathcal F$ are called events.
\end{definition}
Note that $\mathbb P$ is dependent on $\mathcal F$ but not on $\Omega$. We talk about probabilities of \textit{events}, not probabilities of \textit{outcomes}. For example, if you pick a uniform number from the interval $[0, 1]$; then the probability of getting any specific outcome is zero --- but we can define useful events that have non-zero probabilities.

\subsection{Properties of the Probability Measure}
\begin{itemize}
	\item $\prob{\stcomp{A}} = 1 - \prob{A}$, since $A$ and $\stcomp{A}$ are disjoint sets, whose union is $\Omega$
	\item $\prob{\varnothing} = 0$, since it is the complement of $\Omega$
	\item if $A \subseteq B$, then $\prob{A} \leq \prob{B}$
	\item $\prob{A \cup B} = \prob{A} + \prob{B} - \prob{A \cap B}$ using the Inclusion-Exclusion theorem
\end{itemize}

\subsection{Examples of Probability Spaces}
\begin{itemize}
	\item Rolling a fair 6-sided die:
	      \begin{itemize}
		      \item $\Omega = \{ 1, 2, 3, 4, 5, 6 \}$
		      \item $\mathcal F = \mathcal P(\Omega)$
		      \item $\forall \omega \in \Omega, \prob{\{ \omega \}} = \frac{1}{6}$, and if $A \subseteq \Omega$ then $\prob{A} = \frac{\abs{A}}{6}$
	      \end{itemize}

	\item Equally likely outcomes (more generally). Suppose $\Omega$ is some finite set, e.g. $\Omega = \{ \omega_1, \omega_2, \dots, \omega_n \}$. Then we define $\prob{A} = \frac{\abs{A}}{\abs{\Omega}}$. In classical probability, this models picking a random element of $\Omega$.

	\item Picking balls from a bag. Suppose we have $n$ balls with $n$ labels from the set $\{1, \dots, n\}$, indistinguishable by touching. Let us pick $k \leq n$ balls at random from the bag, \textit{without replacement}. Here, `at random' just means that all possible outcomes are equally likely, and their probability measures should be equal.

	      We will take $\Omega = \{ A \subseteq \{1, \dots, n\} : \abs{A} = k \}$. Then $\abs{\Omega} = \binom{n}{k}$. Then of course $\prob{\{ \omega \}} = \frac{1}{\abs*{\Omega}}$, since all outcomes (combinations, in this case) are equally likely.

	\item Consider a well-shuffled deck of 52 cards, i.e. it is equally likely that each possible permutation of the 52 cards will appear. $\Omega = \{ \text{all permutations of 52 cards} \}$, and $\abs*{\Omega} = 52!$

	      The probability that the top two cards are aces is therefore $\frac{4 \times 3 \times 50!}{52!} = \frac{1}{221}$, since there are $4 \times 3 \times 50!$ outcomes that produce such an event.

	\item Consider a string of $n$ random digits from $\{0, \dots, 9\}$. Then $\Omega = \{ 0, \dots, 9 \}^n$, and $\abs*{\Omega} = 10^n$. We define $A_k = \{ \text{no digit exceeds } k \}$, and $B_k = \{ \text{largest digit is } k \}$. Then $\prob{B_k} = \frac{\abs*{B_k}}{\abs*{\Omega}}$. Notice that $B_k = A_k \setminus A_{k-1}$. $\abs*{A_k} = (k+1)^n$, so $\abs*{B_k} = (k+1)^n - k^n$, so $\prob{B_k} = \frac{(k+1)^n - k^n}{10^n}$.

	\item The birthday problem. There are $n$ people; what is the probability that at least two of them share a birthday? We assume that each year has exactly 365 days, i.e. nobody is born on 29\textsuperscript{th} of February, and that the probabilities of being born on any given day are equal.

	      Let $\Omega = \{1, \dots, 365\}^n$, and $\mathcal F = \mathcal P(\Omega)$. Since all outcomes are equally likely, we take $\prob{\{\omega\}} = \frac{1}{365^n}$.

	      Let $A = \{ \text{at least two people share the same birthday} \}$. $\stcomp{A} = \{ \text{all } n \text{ birthdays are different} \}$. Since $\prob{A} = 1 - \prob{\stcomp{A}}$, it suffices to calculate $\prob{\stcomp{A}}$, which is $\frac{\abs*{\stcomp{A}}}{\abs*{\Omega}} = \frac{365!}{(365 - n)!365^n}$. So the answer is $\prob{A} = 1 - \frac{365!}{(365 - n)!365^n}$

	      Note that at $n=22$, $\prob{A} \approx 0.476$ and at $n=23$, $\prob{A} \approx 0.507$. So when there are at least 23 people in a room, the probability that two of them share a birthday is around 50\%.
\end{itemize}

\subsection{Combinatorial Analysis}
Let $\Omega$ be a finite set, and suppose $\abs*{\Omega} = n$. We want to partition $\Omega$ into $k$ disjoint subsets $\Omega_1, \dots, \Omega_k$ with $\abs*{\Omega_i} = n_i$ and $\sum_{i=1}^k n_i = n$. How many ways of doing such a partition are there? The result is
\[ \underbrace{\binom{n}{n_1}}_{\text{choose first set}}\underbrace{\binom{n-n_1}{n_2}}_{\text{choose second set}}\dots\underbrace{\binom{n-(n_1 + n_2 + \dots + n_{k+1})}{n_k}}_{\text{choose last set}} = \frac{n!}{n_1!n_2!\dots n_k!} \]
So we will write
\[ \binom{n}{n_1, \dots, n_k} = \frac{n!}{n_1!n_2!\dots n_k!} \]

Now, let $f\colon \{1, \dots, k\} \to \{1, \dots, n\}$. $f$ is strictly increasing if $x < y \implies f(x) < f(y)$. $f$ is increasing if $x < y \implies f(x) \leq f(y)$. How many strictly increasing functions $f$ exist? Note that if we know the range of $f$, the function is completely determined. The range is a subset of $\{1, \dots, n\}$ of size $k$, i.e. a $k$-subset of an $n$-set, which yields $\binom{n}{k}$ choices, and thus there are $\binom{n}{k}$ strictly increasing functions.

How many increasing functions $f$ exist? Let us define a bijection from the set of increasing functions $\{f\colon \{1, \dots, k\} \to \{1, \dots, n\}\}$ to the set of \textit{strictly} increasing functions $\{g\colon \{1, \dots, k\} \to \{1, \dots, n+k-1\}\}$. For any increasing function $f$, we define $g(i) = f(i) + i - 1$. Then $g$ is clearly strictly increasing, and takes values in the range $\{1, \dots, n+k-1\}$. By extension, we can define an increasing function $f$ from any strictly increasing function $g$. So the total number of increasing functions $f\colon \{1, \dots, k\} \to \{1, \dots, n\}$ is $\binom{n+k-1}{k}$.
