\subsection{Introduction}
Let \( M \) and \( N \) be \( R \)-modules.
Informally, the tensor product of \( M \) and \( N \) over \( R \) is the set \( M \otimes_R N \) of all sums
\[ \sum_{i=1}^\ell m_i \otimes n_i;\quad m_i \in M, n_i \in N \]
subject to the relations
\begin{align*}
    (m_1 + m_2) \otimes n &= m_1 \otimes n + m_2 \otimes n \\
    m \otimes (n_1 + n_2) &= m \otimes n_1 + m \otimes n_2 \\
    (rm) \otimes n &= r(m \otimes n) \\
    m \otimes (rn) &= r(m \otimes n)
\end{align*}
This is a module that abstracts the notion of bilinearity between two modules.
\begin{example}
    Consider \( \faktor{\mathbb Z}{2\mathbb Z} \otimes_{\mathbb Z} \faktor{\mathbb Z}{3\mathbb Z} \).
    In this \( \mathbb Z \)-module,
    \[ x \otimes y = (3x) \otimes y = x \otimes (3y) = x \otimes 0 = x \otimes (0 \cdot 0) = 0 (x \otimes 0) = 0 \]
    Hence \( \faktor{\mathbb Z}{2\mathbb Z} \otimes_{\mathbb Z} \faktor{\mathbb Z}{3\mathbb Z} = 0 \).
\end{example}
\begin{example}
    Now consider \( {\mathbb R}^n \otimes_{\mathbb R} {\mathbb R}^\ell \).
    We will show later that this is isomorphic to \( \mathbb R^{n+\ell} \).
\end{example}

\subsection{Definition and universal property}
\begin{definition}
    A map of \( R \)-modules \( f : M \times N \to L \) is \emph{\( R \)-bilinear} if for each \( m_0 \in M \) and \( n_0 \in N \), the maps \( n \mapsto f(m_0, n) \) and \( m \mapsto f(m, n_0) \) are \( R \)-linear (or equivalently, a homomorphism of \( R \)-modules).
\end{definition}
\begin{definition}
    Let \( M, N \) be \( R \)-modules.
    Let \( \mathcal F = R^{\oplus(M \times N)} \) be the free \( R \)-module with coordinates indexed by \( M \times N \).
    Define \( K \subseteq \mathcal F \) to be the submodule generated by the following set of relations:
    \begin{align*}
        &(m_1 + m_2, n) - (m_1, n) - (m_2, n) \\
        &(m, n_1 + n_2) - (m, n_1) - (m, n_2) \\
        &r (m, n) - (rm, n) \\
        &r (m, n) - (m, rn)
    \end{align*}
    The \emph{tensor product} \( M \otimes_R N \) is \( \faktor{\mathcal F}{K} \).
    We further define the \( R \)-bilinear map
    \[ i_{M \otimes N} : M \times N \to M \otimes N;\quad i_{M \otimes N}(m, n) = m \otimes n \]
\end{definition}
\begin{proposition}[universal property of the tensor product]
    The pair \( (M \otimes_R N, i_{M \otimes_R N}) \) satisfies the following universal property.
    For every \( R \)-module \( L \) and every \( R \)-bilinear map \( f : M \times N \to L \), there exists a unique homomorphism \( h : M \otimes_R N \to L \) such that the following diagram commutes.
    \[\begin{tikzcd}
        {M \times N} & {M \otimes_R N} \\
        & L
        \arrow["{i_{M \otimes_R N}}", from=1-1, to=1-2]
        \arrow["h", dashed, from=1-2, to=2-2]
        \arrow["f"', from=1-1, to=2-2]
    \end{tikzcd}\]
    Equivalently, \( h \circ i_{M \otimes_R N} = f \).
\end{proposition}
\begin{proof}
    The conclusion \( h \circ i_{M \otimes N} = f \) holds if and only if for all \( m, n \), we have
    \[ h(m \otimes n) = f(m, n) \]
    Note that the elements \( \qty{m \otimes n} \) generate \( M \otimes N \) as an \( R \)-module, so there is at most one \( h \).
    We now show that the definition of \( h \) on the pure tensors \( m \otimes n \) extends to an \( R \)-linear map \( M \otimes N \to L \).
    The map \( R^{\oplus(M \otimes N)} \to L \) given by \( (m, n) \mapsto f(m, n) \) exists by the universal property of the direct sum.
    However, this map vanishes on the generators of \( K \), so it factors through the quotient \( \faktor{\mathcal F}{K} \) as required.
\end{proof}
The universal property given above characterises the tensor product up to isomorphism.
\begin{proposition}
    Let \( M, N \) be \( R \)-modules, and \( (T, j) \) be an \( R \)-module and an \( R \)-bilinear map \( M \times N \to T \).
    Suppose that \( (T, j) \) satisfies the same universal property as \( M \otimes N \).
    Then there is a unique isomorphism of \( R \)-modules \( \varphi : M \otimes N \similarrightarrow T \) such that \( \varphi \circ i_{M \otimes N} = j \).
\end{proposition}
\begin{proof}
    By using the universal property of \( M \otimes N \) and \( T \), we obtain \( \varphi \) and \( \psi \) as follows.
    \[\begin{tikzcd}
        {M \otimes N} && T \\
        & {M \times N}
        \arrow["{i_{M \otimes N}}", from=2-2, to=1-1]
        \arrow["j"', from=2-2, to=1-3]
        \arrow["\varphi", shift left, dashed, from=1-1, to=1-3]
        \arrow["\psi", shift left, dashed, from=1-3, to=1-1]
    \end{tikzcd}\]
    The universal property states that \( \varphi \circ i_{M \otimes N} = j \) and \( \psi \circ j = i_{M \otimes N} \).
    Hence, \( \psi \circ \varphi \circ i_{M \otimes N} = i_{M \otimes N} \).
    This means that the following diagram commutes.
    \[\begin{tikzcd}
        {M \times N} && {M \otimes N} \\
        && {M \otimes N}
        \arrow["{i_{M \otimes N}}", from=1-1, to=1-3]
        \arrow["\id"', shift right=2, from=1-3, to=2-3]
        \arrow["{i_{M \otimes N}}"', from=1-1, to=2-3]
        \arrow["{\psi \circ \varphi}", shift left=2, from=1-3, to=2-3]
    \end{tikzcd}\]
    By the uniqueness condition of the universal property, \( \id = \psi \circ \varphi \).
    Similarly, \( \id = \varphi \circ \psi \).
    Hence, \( \varphi \) is an isomorphism \( M \otimes N \to T \) with \( \varphi \circ i_{M \otimes N} = j \).
    Uniqueness of \( \varphi \) is guaranteed by the universal property: it is the only solution to \( \varphi \circ i_{M \otimes N} = j \).
\end{proof}
In particular, we have
\[ \operatorname{Bilin}_R(M \times N, L) \similarrightarrow \Hom(M \otimes_R N, L) \]
given by the universal property, and the inverse is given by \( h \mapsto h \circ i_{M \otimes N} \).

\subsection{Zero tensors}
\begin{proposition}
    Let \( M, N \) be \( R \)-modules.
    Then
    \[ \sum m_i \otimes n_i = 0 \]
    if and only if for every \( R \)-module \( L \) and every \( R \)-bilinear map \( f : M \times N \to L \), we have
    \[ \sum f(m_i, n_i) = 0 \]
\end{proposition}
To show an element of \( M \otimes N \) is nonzero, it suffices to find a single \( R \)-module \( L \) and bilinear map \( M \times N \to L \) with mapping the required sum to a nonzero value.
\begin{proof}
    Assume \( \sum m_i \otimes n_i = 0 \).
    \( f \) factors through the map \( i_{M \otimes N} \), giving
    \[\begin{tikzcd}
        {M \times N} & {M \otimes N} \\
        & L
        \arrow["{i_{M \otimes N}}", from=1-1, to=1-2]
        \arrow["h", from=1-2, to=2-2]
        \arrow["f"', from=1-1, to=2-2]
    \end{tikzcd}\]
    So
    \[ \sum f(m_i, n_i) = \sum h(i_{M \otimes N}(m, n)) = h\qty(\sum i_{M \otimes N} (m, n)) = h(0) = 0 \]
    In the other direction, suppose \( \sum m_i \otimes n_i \neq 0 \).
    Then, \( \sum i_{M \otimes N}(m_i, n_i) \neq 0 \), as required.
\end{proof}
\begin{example}
    Let \( k \) be a field, and consider \( k^m \otimes k^\ell \).
    Let \( k^m \) have basis \( \qty{e_1, \dots, e_m} \) and \( k^\ell \) have basis \( f_1, \dots, f_\ell \).
    Then
    \[ k^m \otimes k^\ell = \vecspan_k \qty{v \otimes w \mid v \in k^m, w \in k^\ell} = \vecspan_k \qty{e_i \otimes f_j} \]
    This is in fact a basis.
    Suppose \( \sum_{i,j} \alpha_{i,j} e_i \otimes f_j = 0 \).
    For each \( a \leq m, b \leq \ell \), define \( T_{a,b} : k^m \times k^\ell \to k \) by
    \[ T_{a,b}\qty((v_i)_{i=1}^k, (w_j)_{j=1}^\ell) = v_a w_b \]
    By the above proposition,
    \[ 0 = \sum_{i,j} \alpha_{i,j} T_{a,b}(e_i, f_j) = \alpha_{a,b} \]
    So \( k^m \otimes k^\ell \simeq k^{m\ell} \).
    Note that this construction only relied on the existence of a free basis, not on \( k \) being a field.
\end{example}
\begin{example}
    Consider \( \mathbb R^2 \otimes_{\mathbb R} \mathbb R^2 \).
    There are infinitely many pure tensors, but there is a basis consisting of the four pure vectors
    \[ e_1 \otimes f_1;\quad e_1 \otimes f_2;\quad e_2 \otimes f_1;\quad e_2 \otimes f_2 \]
    A pure tensor in \( \mathbb R^2 \otimes_{\mathbb R} \mathbb R^2 \) is of the form
    \[ (\alpha e_1 + \beta e_2) \otimes (\gamma f_1 + \delta f_2) \]
    which expands to
    \[ (\alpha\gamma)(e_1 \otimes f_1) + (\alpha\delta)(e_1 \otimes f_2) + (\beta\gamma)(e_2 \otimes f_1) + (\beta\delta)(e_2 \otimes f_2) \]
    Note that there is a linear dependence relation between the coefficients \( \alpha\gamma, \alpha\delta, \beta\gamma, \beta\delta \), so in some sense `most' tensors are not pure.
    For example,
    \[ 1(e_1 \otimes f_1) + 2(e_1 \otimes f_2) + 3(e_2 \otimes f_1) + 4(e_2 \otimes f_2) \]
    is not pure.
\end{example}
\begin{example}
    Consider \( \mathbb Z \otimes_{\mathbb Z} \faktor{\mathbb Z}{2\mathbb Z} \).
    In this module,
    \[ 2 \otimes (1 + 2\mathbb Z) = 1 \otimes (2 + 2\mathbb Z) = 1 \otimes 0 = 0 \]
    Note that \( \mathbb Z \) has a \( \mathbb Z \)-submodule \( 2\mathbb Z \).
    In \( 2\mathbb Z \otimes_{\mathbb Z} \faktor{\mathbb Z}{2\mathbb Z} \), the element also denoted with \( 2 \otimes (1 + 2\mathbb Z) \) is nonzero.
    For example, we can define a bilinear map to \( \faktor{\mathbb Z}{2\mathbb Z} \) given by
    \[ b(2n, x + 2\mathbb Z) = n x + 2\mathbb Z \]
    Then \( b(2, 1 + 2\mathbb Z) = 1 \neq 0 \).
    So it is not the case that tensor products of submodules are submodules of tensor products.

    However, if \( M' \subseteq M \) and \( N' \subseteq N \) and \( \sum m_i \otimes n_i = 0 \) in \( M' \otimes N' \), then \( \sum m_i \otimes n_i = 0 \) in \( M \otimes N \).
\end{example}
\begin{proposition}
    If \( \sum m_i \otimes n_i = 0 \) in \( M \otimes_R N \), then there are finitely generated \( R \)-submodules \( M' \subseteq M \) and \( N' \subseteq N \) such that the expression \( \sum m_i \otimes n_i \) also evaluates to zero in \( M' \otimes_R N' \).
\end{proposition}
This is the last proof that will use the direct construction of the tensor product instead of the universal property directly.
\begin{proof}
    We know that \( \sum m_i \otimes n_i = 0 \) in \( M \otimes_R N = \faktor{R^{\oplus(M \times N)}}{K} \), so in particular \( \sum e_{(m_i, n_i)} \in K \), where \( e_x \) maps \( x \in M \times N \) to its basis element in \( R^{\oplus(M \times N)} \).
    So this is a finite sum of \( \alpha_i k_i \) with \( \alpha_i \in R, k_i \in K \), and so we can take the \( m_1', \dots, m_a' \) that appear on the left-hand sides of the \( k_i \) as the generators for \( M' \), and similarly for \( N' \).
\end{proof}
\begin{corollary}
    Let \( A, B \) be torsion-free abelian groups.
    Then \( A \otimes_{\mathbb Z} B \) is torsion-free.
\end{corollary}
\begin{proof}
    Suppose \( n \qty(\sum a_i \otimes b_i) = 0 \) with \( n \geq 1 \).
    By the previous proposition, there are finitely generated subgroups \( A' \leq A \) and \( B' \leq B \) such that \( n \qty(\sum a_i \otimes b_i) = 0 \) in \( A' \otimes_{\mathbb Z} B' \).
    But as \( A' \) and \( B' \) are finitely generated abelian groups, the structure theorem shows that \( A' = \mathbb Z^m \) and \( B' = \mathbb Z^\ell \), showing that \( A' \otimes_{\mathbb Z} B' \simeq \mathbb Z^{m\ell} \) is torsion-free.
    Thus \( \sum a_i \otimes b_i = 0 \) in \( A' \otimes_{\mathbb Z} B' \), so also \( \sum a_i \otimes b_i = 0 \) in \( A \otimes_{\mathbb Z} B \).
\end{proof}
\begin{example}
    \[ \mathbb C^2 \otimes_{\mathbb C} \mathbb C^3 \simeq \mathbb C^6 \simeq \mathbb R^{12} \]
    However,
    \[ \mathbb C^2 \otimes_{\mathbb R} \mathbb C^3 \simeq \mathbb R^4 \otimes_{\mathbb R} \mathbb R^6 \simeq \mathbb R^{24} \]
    This is to be expected: tensoring over a larger ring introduces more relations, so the amount of distinguishable elements should shrink.
\end{example}

\subsection{Monoidal structure}
\begin{proposition}[commutativity]
    There is an isomorphism \( M \otimes N \simeq N \otimes N \) mapping a pure tensor \( m \otimes n \) to \( n \otimes m \).
\end{proposition}
\begin{proposition}[associativity]
    There is an isomorphism \( (M \otimes N) \otimes P \simeq M \otimes (N \otimes P) \) mapping \( (m \otimes n) \otimes p \) to \( m \otimes (n \otimes p) \).
\end{proposition}
% to solve, define tensor product of three modules
\begin{proposition}[identity]
    There is an isomorphism \( R \otimes M \simeq M \) mapping \( r \otimes m \) to \( rm \).
\end{proposition}
\begin{proposition}[distributivity]
    There is an isomorphism \( \qty(\bigoplus_i M_i) \otimes P \simeq \bigoplus_i (M_i \otimes P) \) mapping % TODO
\end{proposition}
\begin{example}
    \[ R^m \otimes_R R^\ell = \qty(\bigoplus_{i=1}^m R) \otimes_R \qty(\bigoplus_{j=1}^\ell R) \simeq \bigoplus_{i=1}^m \bigoplus_{j=1}^\ell (R \otimes R) \simeq \bigoplus_{i=1}^m \bigoplus_{j=1}^\ell R \simeq R^{m\ell} \]
\end{example}

\subsection{Tensor products of maps}
\begin{proposition}
    Let \( f : M \to M' \) and \( g : N \to N' \) be \( R \)-module homomorphisms.
    There is a unique \( R \)-module homomorphism \( f \otimes g : M \otimes N \to M' \otimes N' \) such that
    \[ (f \otimes g)(m \otimes n) = f(m) \otimes g(n) \]
\end{proposition}
\begin{proof}
    We apply the universal property to the map \( T : M \times N \to M \otimes N' \) given by
    \[ T(m, n) = f(m) \otimes g(n) \]
    which can be checked to be \( R \)-bilinear.
\end{proof}
