\subsection{Sinusoidal Forcing}
Let
\[ \ddot y + \mu \dot y + \omega_0^2 y = \sin \omega t \]
Now let us guess $y_p = A\sin \omega t + B\cos \omega t$. Equating coefficients of $\sin\omega t$ gives
\[ -A\omega^2 - B\mu\omega + \omega_0^2A = 1 \tag{a} \]
and equating $\cos\omega t$ gives
\[ -B\omega^2 + A\mu\omega + \omega_0^2B = 0 \tag{b} \]
Then:
\[ (\text b) \implies A = B \frac{\omega^2 - \omega_0^2}{\mu\omega} \]
\[ (\text a) \implies A(\omega_0^2 - \omega^2) = 1 + B\mu\omega \]
So
\[ A = \frac{\omega_0^2 - \omega^2}{(\omega_0^2 - \omega^2)^2 + \mu^2\omega^2} \]
\[ B = \frac{-\mu\omega}{(\omega_0^2 - \omega^2)^2 + \mu^2\omega^2} \]
Altogether, we have
\[ y_p = \frac{1}{(\omega_0^2 - \omega^2)^2 + \mu^2\omega^2}\left[ (\omega_0^2 - \omega^2)\sin\omega t - \mu \omega \cos \omega t \right] \]
Drawing an example of this kind of particular integral in red (with the complementary function in blue), we can see the following:\medskip

\begin{tikzpicture}
	\begin{axis}[
			%axis lines = left,
			xlabel = $t$,
			ylabel = $y$,% $y = y_c + y_p$,
			width=15cm,
			height=5cm,
			xmin=0,
			xmax=10,
			ymin=-2.4,
			ymax=2.4
			%xticklabel=\empty,
			%yticklabel=\empty
		]

		\addplot [
			domain=0:10,
			samples=200,
			color=blue
		]
		{2.5*e^(-0.4*x)*sin(3*deg(x))};

		\addplot [
			domain=0:10,
			samples=200,
			color=red
		]
		{0.3*(sin(7*deg(x)) + cos(7*deg(x)))};
	\end{axis}
\end{tikzpicture}

\noindent And adding both together to form a particular solution gives:\medskip

\begin{tikzpicture}
	\begin{axis}[
			%axis lines = left,
			xlabel = $t$,
			ylabel = $y$,% $y = y_c + y_p$,
			width=15cm,
			height=5cm,
			xmin=0,
			xmax=10,
			ymin=-2.4,
			ymax=2.4
			%xticklabel=\empty,
			%yticklabel=\empty
		]

		\addplot [
			domain=0:10,
			samples=200,
			color=black
		]
		{2.5*e^(-0.4*x)*sin(3*deg(x)) + 0.3*(sin(7*deg(x)) + cos(7*deg(x)))};
	\end{axis}
\end{tikzpicture}

\noindent Let us make a few comments about these forced oscillations.
\begin{itemize}
	\item The complementary function gives us the transient (short-term) response to the initial conditions.
	\item The particular integral gives the long-term response to the forcing term.
	\item In some sense, the system `forgets' about the initial conditions over time due to the damping term.
\end{itemize}

\subsection{Resonance in Undamped Systems}
What happens if $\omega = \omega_0$? If $\mu \neq 0$ (i.e. it is a damped system), then
\[ \lim_{\omega \to \omega_0} y_p = \frac{-\cos\omega_0 t}{\mu\omega_0} \]
This is a finite amplitude oscillation. Note that the amplitude increases with decreasing $\mu$, so clearly this solution has a problem at $\mu = 0$. To work with this, we'll let
\[ \ddot y + \omega_0^2 y = \sin\omega_0 t \]
We will use detuning to get solutions for this equation. Consider instead
\[ \ddot y + \omega_0^2 y = \sin\omega t \]
where $\omega \neq \omega_0$. We will guess that the particular integral is of the form $y_p = C\sin\omega t$ since by inspection there cannot be any cosine terms.
\[ C(-\omega^2 + \omega_0^2) = 1 \]
\[ \therefore y_p = \frac{1}{\omega_0^2 - \omega^2}\sin\omega t \]
As the system is linear in $y$ and its derivatives, we can freely add some multiple of the complementary function and it will remain a solution.
\[ y_p = \frac{1}{\omega_0^2 - \omega^2}\sin\omega t + A \sin\omega_0 t \]
Now let us pick $A = \frac{-1}{\omega_0^2 - \omega^2}$, so
\[ y_p = \frac{\sin \omega t - \sin \omega_0 t}{\omega_0^2 - \omega^2} \]
Rewriting this using angle addition and subtraction identities:
\[ y_p = \frac{2}{\omega_0^2 - \omega^2}\left[ \cos\left( \frac{\omega + \omega_0}{2}t \right) \sin\left( \frac{\omega - \omega_0}{2}t \right) \right] \]
For convenience, let $\Delta\omega \equiv \omega_0 - \omega$, and therefore $\frac{\omega + \omega_0}{2} = w_0 - \frac{1}{2}\Delta\omega$.
\[ y_p = \frac{-2}{\Delta\omega(\omega_0 + \omega)}\left[ \cos\left( \left(\omega_0 - \frac{\Delta\omega}{2}\right)t \right) \sin\frac{\Delta\omega t}{2} \right] \]
In the following diagram, $y_p$ is drawn in blue, with the sine term (in red) acting as an envelope for the higher-frequency cosine term. The phenomenon visible here is known as `beating', as an oscillator with a fundamental frequency slightly different to the forcing frequency will begin oscillating then stop, and repeat this cycle.\medskip

\begin{tikzpicture}
	\begin{axis}[
			%axis lines = left,
			xlabel = $t$,
			ylabel = $y_p$,% $y = y_c + y_p$,
			width=15cm,
			height=5cm,
			xmin=0,
			xmax=10,
			ymin=-2.4,
			ymax=2.4
			%xticklabel=\empty,
			%yticklabel=\empty
		]

		\addplot [
			domain=0:10,
			samples=200,
			color=red
		]
		{sin(deg(x))};
		\addplot [
			dashed,
			domain=0:10,
			samples=200,
			color=red
		]
		{-sin(deg(x))};

		\addplot [
			domain=0:10,
			samples=2000,
			color=blue
		]
		{cos(24.3*deg(x))*sin(deg(x))};
	\end{axis}
\end{tikzpicture}

\noindent As we reduce $\Delta\omega$ to zero, we have
\[ \lim_{\Delta\omega\to 0} \sin\left(\frac{\Delta\omega}{2}t\right) \approx \frac{\Delta\omega}{2}t \]
So
\begin{align*}
	\lim_{\Delta\omega\to 0}
	 & \approx \frac{-2}{\Delta\omega(\omega_0 + \omega_0)} \cos(\omega_0t)\left(\frac{\Delta\omega}{2}t\right) \\
	 & \approx \frac{-2t}{\omega_0} \cos\omega_0t                                                               \\
\end{align*}
This is linear growth in amplitude over time. This increase is unbounded an in undamped system. Note that $y_p$ takes the form of one of the complementary functions multiplied by the independent variable.

\subsection{Impulses and Point Forces}
Consider a system that experiences a sudden force, for example a car's suspension when driving over a speed bump. Let us define $y$ to be the displacement from the undisturbed height of the suspension. Let the car's mass be $M$. In a small finite window $\varepsilon$ around some time $T$, the excess force $F$ (the forcing term) on the system is greater than zero. As $\varepsilon$ tends to zero, the force becomes a sudden impulse. Let us model this using the equation
\[ M\ddot y = F(t) - ky - L\dot y \]
We can see that before time $T$, $y=0$. After this point, there is some kind of oscillation. Note that the value of $y$ is always continuous (otherwise this would violate many laws of physics), but the derivative is not necessarily continuous at the point $T$. Let us integrate the equation above in time from $T - \varepsilon$ to $T + \varepsilon$.
\begin{equation}\label{impulse1}
	\lim_{\varepsilon \to 0} \left[ M[\dot y]_{T - \varepsilon}^{T + \varepsilon} = \int_{T - \varepsilon}^{T + \varepsilon} F(t) \dd{t} - k \underbrace{\int_{T - \varepsilon}^{T + \varepsilon} y \dd{t}}_{0\text{ if $y$ is finite}} - L\underbrace{[y]_{T - \varepsilon}^{T + \varepsilon}}_{0\text{ if $y$ is continuous}} \right]
\end{equation}
We now can define the impulse $I$ to be
\[ I = \lim_{\varepsilon \to 0} \int_{T - \varepsilon}^{T + \varepsilon} F(t) \dd{t} \]
Hence
\[ \eqref{impulse1} \implies I = \lim_{\varepsilon \to 0} M[\dot y]_{T - \varepsilon}^{T + \varepsilon} \]
So if the impulse is nonzero, the velocity $\dot y$ experiences a sudden change, so it is discontinuous at $T$. The value of this sudden change in velocity depends on the integral of the force.
