\subsection{Intuition with geometry}
Which is more symmetrical, a scalene triangle or an equilateral triangle?
Clearly, the equilateral triangle has more symmetries; you can rotate it \(120\degree\) or \(240\degree\), and you can reflect it across three axes.
The scalene triangle has no symmetries that modify the object, but by convention we call the `do-nothing' operation a symmetry as well.

By a `symmetry' of an object, we mean something that we can do to it that preserves its structure.
In the case of these shapes, we want to preserve the vertices and edges; these symmetries are rotations and reflections.
For the equilateral triangle then, what are all the symmetries?

\begin{tikzpicture}
	\draw (0,0) node[below]{\(C\)}
	-- (2,0) node[below]{\(B\)}
	-- (1,1.73) node[above]{\(A\)}
	-- cycle;
	\draw(1, 0.6) node{\(e\)};

	\draw (2.5,0) node[below]{\(B\)}
	-- (4.5,0) node[below]{\(C\)}
	-- (3.5,1.73) node[above]{\(A\)}
	-- cycle;
	\draw(3.5, 0.6) node{\(s\)};

	\draw (5,0) node[below]{\(B\)}
	-- (7,0) node[below]{\(A\)}
	-- (6,1.73) node[above]{\(C\)}
	-- cycle;
	\draw(6, 0.6) node{\(r\)};

	\draw (7.5,0) node[below]{\(A\)}
	-- (9.5,0) node[below]{\(C\)}
	-- (8.5,1.73) node[above]{\(B\)}
	-- cycle;
	\draw(8.5, 0.6) node{\(r^2\)};

	\draw (10,0) node[below]{\(A\)}
	-- (12,0) node[below]{\(B\)}
	-- (11,1.73) node[above]{\(C\)}
	-- cycle;
	\draw(11, 0.6) node{\(sr\)};

	\draw (12.5,0) node[below]{\(C\)}
	-- (14.5,0) node[below]{\(A\)}
	-- (13.5,1.73) node[above]{\(B\)}
	-- cycle;
	\draw(13.5, 0.6) node{\(sr^2\)};
\end{tikzpicture}

As stated before, we assign the letter \(e\) to the identity element.
The operation \(s\) is a reflection; \(r\) is a rotation.
By combining these elements, we get the set of elements of the group.
Note that order matters: \(sr\neq rs\).

\subsection{Basic notion}
\begin{definition}[Group]
	A group is a set \(G\) together with a way of composing its elements \(\ast\) satisfying (\(\forall g, h, k \in G\)):
	\begin{itemize}
		\item (closure) \(g \ast h \in G\)
		\item (identity) \(\exists e \in G \mathrm{\ s.t.\ } e \ast g = g \ast e = g\)
		\item (inverses) \(\exists g^{-1} \in G \mathrm{\ s.t.\ } g \ast g^{-1} = g^{-1} \ast g = e\)
		\item (associativity) \(g \ast (h \ast k) = (g \ast h) \ast k\)
	\end{itemize}
\end{definition}

Formally, we might say that a set \(G\) with a binary operation \(\ast : G \times G \to G\) is a group if it follows the last three axioms; the first rule is implicit in the function's type.

\subsection{Examples}
\begin{enumerate}
	\setcounter{enumi}{-1}
	\item \(G = \{ e \}\) --- this is the `trivial group'.
	\item \(G = \{ \text{symmetries of the equilateral triangle} \} \); \(\ast\) is defined by: `\(g \ast h\) means doing \(h\) then \(g\)'.
	\item \(G = (\mathbb Z, +)\).
	      This is easy to prove by verifying the axioms.
	\item \(G = (\mathbb R, +); (\mathbb Q, +); (\mathbb C, +)\)
	\item \(G = (\mathbb R^*, \cdot)\) where \(\mathbb R^* = \mathbb R \setminus \{ 0\}\).
	      Note that \((\mathbb R, \cdot)\) is not a group, because \(\nexists 0^{-1} \in \mathbb R \st 0^{-1} \cdot 0 = 0 \cdot 0^{-1} = 1\).
	\item \(G = (\mathbb R, \ast)\) where \(r \ast s := r + s + 5\).
	\item \(G = (\mathbb Z_n, +)\) where \(\mathbb Z_n = \{ 0, 1, 2, \cdots, n-1 \}\) and addition is done modulo \(n\).
	\item A vector space with the operation of vector addition is a group.
	\item \(GL_2(\mathbb R)\) is the set of invertible \(2\times 2\) matrices, which is a group with respect to matrix multiplication.
\end{enumerate}

\subsection{Non-examples}
\begin{enumerate}
	\item \(G = (\mathbb Z_n, +)\) where addition is not performed modulo \(n\).
	      This group is not closed, e.g.
	      \((n-1) + 2 \notin G\).
	\item \(G = (\mathbb Z, \cdot)\) because \(\nexists n\in \mathbb Z \st 2 \cdot n = n \cdot 2 = 1\).
	\item \(G = (\mathbb R, \ast)\) where \(r \ast s = r^2 s\) because there is no identity element.
	\item \(G = (\mathbb N, \ast)\) where \(n \ast m := \abs{n - m}\) because it is non-associative, e.g.
	      \(1 \ast (2 \ast 5) = 2\); \((1 \ast 2) \ast 5 = 4\).
\end{enumerate}

\subsection{Notation}
We use the notation \(gh=g\cdot h = g \ast h\) here to represent the group operation (regardless of the specific operation in question).
\subsection{Basic group properties}
\begin{proposition}
	Let \(G\) be a group.
	Then,
	\begin{enumerate}[i.]
		\item The identity element \(e\) is unique.
		\item \(\forall g \in G\), the inverse \(g^{-1}\) is unique.
		\item \(g \cdot h = g \iff h \cdot g = g\)
		\item \(g \cdot h = e \iff h \cdot g = e\)
		\item \((gh)^{-1} = h^{-1} g^{-1}\)
		\item \((g^{-1})^{-1} = g\)
	\end{enumerate}
\end{proposition}
\begin{proof}
	We prove each case individually.
	\begin{enumerate}[i.]
		\item Assume \(\exists e, e'\) which are distinct identity elements.
		      We have \(e e' = e\) and \(e e' = e'\) by the definition of the inverse so \(e = e'\) \contradiction{}
		\item Suppose \(h\) and \(k\) are distinct inverses of \(g\).
		      Then \(gh = e\) and \(gk = e\), so:
		      \begin{align*}
			      gh        & = gk                       \\
			      g^{-1} gh & = g^{-1} gk                \\
			      h         & = k \text{ \contradiction}
		      \end{align*}
		\item \begin{align*}
			      gh      & = g  \\
			      \iff gh & = ge \\
			      \iff h  & = e  \\
			      \iff hg & = eg \\
			      \iff hg & = g
		      \end{align*}
		\item \begin{align*}
			      gh              & = e       \\
			      \iff ghg        & = g       \\
			      \iff g^{-1} ghg & = g^{-1}g \\
			      \iff hg         & = e
		      \end{align*}
		\item \((gh) (h^{-1}g^{-1}) = g h h^{-1} g^{-1} = g g^{-1} = e\)
		\item \(g^{-1} g = e\)
	\end{enumerate}
\end{proof}

\begin{definition}[abelian group]
	A group \(G\) is said to be \textit{abelian} if \(\forall a, b, \in G\), \(a \ast b = b \ast a\).
\end{definition}
A common example of an abelian group is the reals under addition.
A non-example is the group of invertible \(2\times 2\) matrices under matrix multiplication.

\begin{definition}
	The order of a group \(G\), denoted \(\abs{G}\), is the number of elements in the set \(G\).
	A group \(G\) is called a finite group if its order is finite, and it is called an infinite group if its order is infinite.
\end{definition}

\subsection{Subgroups}
\begin{definition}
	Let \((G, \ast)\) be a group.
	A subset \(H \subseteq G\) is a subgroup of \(G\) if \((H, \ast)\) is a group.
	We denote this \(H \leq G\).
\end{definition}
We must verify each group axiom on a subset to check if it is a subgroup --- with the notable exception of the associativity axiom, the property of associativity is inherited by subgroups.
Here are some examples of subgroups.

\begin{enumerate}
	\setcounter{enumi}{-1}
	\item \(\{ e \}\) is the trivial subgroup
	\item \(G \leq G\)
	\item \((\mathbb Z, +) \leq (\mathbb Q, +) \leq (\mathbb R, +) \leq (\mathbb C, +)\)
\end{enumerate}
\begin{lemma}
	Let \(G\) be a group.
	\(H \subset G\) is a subgroup of \(G\) if and only if \(H\) is non-empty and \(\forall a, b \in H, ab^{-1} \in H\).
\end{lemma}
\begin{proof}
	We prove each axiom.
	\begin{itemize}
		\item (identity) Setting \(a = b\) gives \(a a^{-1} = e \in H\) as required.
		\item (inverses) Setting \(a = e\), which we know exists from the identity proof above, gives \(b^{-1} \in H\).
		\item (closure) Setting \(b = c^{-1}\), we know that \(c \in H\), and we can always choose a \(b\) such that \(c\) is any value we want; and with the property we can see that \(ac \in H\) as required by the closure axiom.
	\end{itemize}
\end{proof}

\begin{proposition}
	The subgroups of \((\mathbb Z, +)\) are precisely the subsets of the form \(n\mathbb Z \subset Z\) where \(n\mathbb Z := \{nk : k\in \mathbb Z\}\).
\end{proposition}
\begin{proof}
	First, we know that each \(n\mathbb Z\) is a subgroup: given any integer \(n \in \mathbb N\) the axioms hold:
	\begin{itemize}
		\item (closure) given \(nk_1, nk_2 \in n\mathbb Z\), we have \(nk_1 + nk_2 = n(k_1 + k_2) \in n\mathbb Z\)
		\item (identity) \(e = 0 = n \cdot 0 \in n\mathbb Z\)
		\item (inverse) \(-nk = n(-k) \in n\mathbb Z\)
	\end{itemize}
	We also prove the converse statement, namely that the only viable subgroups are of the form \((n\mathbb Z, +)\).
	If \(H = \{ 0 \}\) then clearly \(H = 0\mathbb Z\) which is a trivial subgroup.
	Otherwise, there are some non-zero elements.

	There must be at least one positive element in \(H\), since any negative element can be inverted to make a positive one in \(H\).
	So, let the smallest positive element be \(n\).
	Since \(H\) is a subgroup, it is closed and has inverses.
	This implies that
	\begin{align*}
		n+n+n+\cdots                & \in H; \\
		n^{-1}+n^{-1}+n^{-1}+\cdots & \in H
	\end{align*}
	Therefore \(n\mathbb Z\) is contained within \(H\).
	Now, let us show that there are no extra elements.
	Suppose, for purposes of a contradiction, that \(\exists k \in H \st k \notin n\mathbb Z\).
	Then, since \(k\) is an integer and not a multiple of \(n\), it must lie between two such multiples: \(nm < k < n(m+1)\) where \(m \in \mathbb Z\).
	This means that \(0 < k - nm < n\) which implies that there is a smaller positive element than \(n\) in the set.
	This is a contradiction, so there are no more elements in the set.
\end{proof}

\begin{proposition}
	The following statements are true:
	\begin{itemize}
		\item Let \(H, K\) be subgroups of \(G\).
		      Then \(H \cap K\) is a subgroup of \(G\).
		\item \(K \leq H\) and \(H \leq G\), then \(K \leq G\).
		\item If \(K \subseteq H\), \(H \leq G\) and \(K \leq G\) then \(K \leq H\).
	\end{itemize}
\end{proposition}
% TODO proofs?
% especially for first part

\subsection{Lattice of subgroups}
We can use a lattice diagram to denote subgroups.
Points below other points joined by lines represent subgroups.
Let \(G, H, K\) be groups and \(H \leq G\) and \(K \leq G\).

\begin{tikzpicture}
	\draw (0,0) node[below]{\(\{e\}\)}
	-- (0,1) node[right]{\(H \cap K\)}
	-- (1,2) node[right]{\(K\)}
	-- (0,3) node[above]{\(G\)}
	-- (-1,2) node[left]{\(H\)}
	-- (0, 1);
	\filldraw (0,0) circle(1.5pt);
	\filldraw (0,1) circle(1.5pt);
	\filldraw (0,3) circle(1.5pt);
	\filldraw (1,2) circle(1.5pt);
	\filldraw (-1,2) circle(1.5pt);
\end{tikzpicture}

\begin{definition}
	Let \(X \neq \varnothing\) be a subset of a group \(G\).
	The subgroup \textit{generated by} \(X\) denoted \(\genset{X}\) is the intersection of all subgroups containing \(X\).
	Equivalently, \(\genset{X}\) is the smallest subgroup of \(G\) that contains \(X\) as a subset.
	Note that there will always exist some subgroup \(\genset{X}\) regardless of what \(X\) is chosen; a trivial result would be \(G\) itself.
\end{definition}

\subsection{Properties of generated groups}
We can make a more precise definition of generated groups as follows:
\begin{itemize}
	\item \(\genset{X}\) contains \(e\)
	\item \(\genset{X}\) contains the set \(X\)
	\item \(\genset{X}\) contains all possible products of \(X\) and their inverses
\end{itemize}

\begin{proposition}
	Let \(X \subseteq G, X \neq \varnothing\).
	Then \(\genset{X}\) is the set of elements of \(G\) of the form \(x_1^{\alpha_1} x_2^{\alpha_2} x_3^{\alpha_3} \cdots x_k^{\alpha_k}\) where \(x_i \in X\) (not necessarily distinct), \(\alpha_i = \pm 1\), and \(k \geq 0\).
	By convention, the empty product \(k=0\) is defined to be \(e\).
\end{proposition}
\begin{proof}
	Let \(T\) be the set of such elements of the given form.
	Clearly, \(T \subseteq \genset{X}\).
	Also, \(T\) is a subgroup of \(G\), and \(X \subseteq T\), so \(\genset{X} \subseteq T\).
	Because both \(T \subseteq \genset{X}\) and \(\genset{X} \subseteq T\), we have \(T = \genset{X}\).
\end{proof}
Note that generating sets are not necessarily unique.
For example, the group of integers under addition generated by \(\genset{1}\) is equivalent to \(\genset{2, 3}\), both of which are equivalent to \(\mathbb Z\), for example.
As a discrete example, \(\mathbb Z_5\) can be generated by any element in the set apart from zero, for example: \(\mathbb Z_5 = \genset{1} = \genset{2} = \genset{3} = \genset{4} \neq \genset{0}\).
