\subsection{Normal Subgroups with Conjugation}
\begin{proposition}
	Normal subgroups are those subgroups that are unions of conjugacy classes.
	Recall that \(\ccl(h) = \{ ghg^{-1} : g \in G \}\).
\end{proposition}
\begin{proof}
	Let \(M \trianglelefteq G\).
	Then if \(h \in N\), then \(ghg^{-1} \in N\) for all \(g \in G\) because \(N\) is a normal subgroup.
	So \(\ccl(h) \subseteq N\).
	So \(N\) is a union of conjugacy classes of its elements;
	\[
		N = \bigcup_{h \in N} \ccl(h)
	\]
	Conversely, if \(H\) is a subgroup that is a union of conjugacy classes, then \(\forall g \in G, \forall h \in H\), we have \(ghg^{-1} \in H\).
	So \(H \trianglelefteq G\).
\end{proof}
As an example, consider \(A_3 = \{ e, (1\ 2\ 3), (1\ 3\ 2) \} \trianglelefteq S_3\).
Now, \(A_3 = \{ e \} \sqcup \{ (1\ 2\ 3), (1\ 3\ 2) \}\).
Note that \((1\ 2\ 3), (1\ 3\ 2)\) are conjugates in \(S_3\) but they are not conjugates in \(A_3\).

\subsection{Conjugation in Symmetric Groups}
\begin{lemma}
	Given a \(k\)-cycle \((a_1\dots a_k)\) and \(\sigma \in S_n\), we have
	\[
		\sigma (a_1\dots a_k) \sigma^{-1} = (\sigma(a_1)\dots \sigma(a_k))
	\]
\end{lemma}
\begin{proof}
	Let us apply the left hand side transformation to \(\sigma(a_i)\).
	\[
		\sigma (a_1\dots a_k) \sigma^{-1} \sigma(a_i) = \sigma (a_1\dots a_k) (a_i) = \sigma(a_{i+1\mod k})
	\]
	Now let us consider the effect of the transformation on \(\sigma(b)\) for \(b \neq a_i\).
	\[
		\sigma (a_1\dots a_k) \sigma^{-1} \sigma(b) = \sigma (a_1\dots a_k) (b) = \sigma(b)
	\]
	So these are unchanged.
	Therefore, the left hand side is equal to the right hand side.
\end{proof}

\begin{proposition}
	Two elements of \(S_n\) are conjugate (in \(S_n\), i.e.\ via a conjugation by some element in \(S_n\)) if and only if they have the same cycle type.
\end{proposition}
\begin{proof}
	Two elements that are conjugate will have the same cycle type: given \(\sigma \in S_n\), we can write \(\sigma\) as a product of disjoint cycles, say \(\sigma = \sigma_1\dots\sigma_m\).
	Then if \(\rho \in S_n\), \(\rho \sigma \rho^{-1} = \rho \sigma_1 \rho^{-1} \rho \sigma_2 \rho^{-1} \dots \rho \sigma_m \rho^{-1}\) which is a product of the conjugates of the cycles.
	By the above lemma, the conjugate of a \(k\)-cycle is a \(k\)-cycle, and because \(\rho\) is bijective the \(\rho \sigma_i \rho^{-1}\) are all disjoint, so we retain the cycle type of \(\sigma\) under conjugation in \(S_n\).

	Conversely, if \(\sigma\) and \(\tau\) have the same cycle type, then we can write
	\[
		\sigma = (a_1\dots a_{k_1})(a_{k_1+1}\dots a_{k_2})\dots
	\]
	\[
		\tau = (b_1\dots b_{k_1})(b_{k_1+1}\dots b_{k_2})\dots
	\]
	in disjoint cycle notation, including singletons.
	Then all of \(\{ 1, \dots, n \}\) appear in both \(\sigma\) and \(\tau\).
	Then, setting \(\rho\) to be defined by \(\rho(a_i) = b_i\), which is indeed a permutation, we obtain \(\rho\sigma\rho^{-1}=\tau\).
\end{proof}
Let us consider the conjugacy classes of \(S_4\).
We can compute the size of \(C_{S_4}\) using the orbit-stabiliser theorem; the conjugacy class is the orbit of a particular point under conjugation.\medskip

\noindent\begin{tabular}{ccccc}
	cycle type  & example element  & size of ccl & size of \(C_{S_4}\) & sign   \\\midrule
	\(1,1,1,1\) & \(e\)            & 1           & 24                  & \(+1\) \\
	\(2,1,1\)   & \((1\ 2)\)       & 6           & 4                   & \(-1\) \\
	\(2,2\)     & \((1\ 2)(3\ 4)\) & 3           & 8                   & \(+1\) \\
	\(3,1\)     & \((1\ 2\ 3)\)    & 8           & 3                   & \(+1\) \\
	\(4\)       & \((1\ 2\ 3\ 4)\) & 6           & 4                   & \(-1\) \\
\end{tabular}

From this, we can compute all normal subgroups of \(S_4\), since normal subgroups:
\begin{itemize}
	\item must contain \(e\)
	\item must be a union of conjugacy classes
	\item must have an order that divides \(\abs{S_4} = 24\)
\end{itemize}
To check all possibilities, we will look through all divisors of \(24\), and check whether we can form a union of conjugacy classes.
\begin{itemize}
	\item (1) \(\{ e \}\)
	\item (2) impossible, no conjugacy classes have orders which add to 2
	\item (3) impossible
	\item (4) \(3+1=4\) so we have
	      \[
		      \{ e, (1\ 2)(3\ 4), (1\ 3)(2\ 4), (1\ 4)(2\ 3) \} \cong C_2 \times C_2
	      \]
	      This subgroup is often referred to as \(V_4\), the Klein four group.
	\item (6) impossilbe
	\item (8) impossible
	\item (12) \(1+3+8=12\) so we have
	      \[
		      \{ e, (1\ 2)(3\ 4), (1\ 3)(2\ 4), (1\ 4)(2\ 3), (1\ 2\ 3), (1\ 3\ 2), (1\ 2\ 4), (1\ 4\ 2), (1\ 3\ 4), (1\ 4\ 3), (2\ 3\ 4), (2\ 4\ 3) \} = A_4
	      \]
	\item (24) \(S_4 \trianglelefteq S_4\).
\end{itemize}
So all possible quotients of \(S_4\) are:
\begin{itemize}
	\item \(\frac{S_4}{\{ e \}} \cong S_4\)
	\item \(\frac{S_4}{V_4} = \{ V_4, (1\ 2)V_4, (1\ 3)V_4, (2\ 3)V_4, (1\ 2\ 3)V_4, (1\ 3\ 2)V_4 \} \cong S_3\)
	\item \(\frac{S_4}{A_4} \cong C_2\)
	\item \(\frac{S_4}{S_4} \cong \{ e \}\)
\end{itemize}
Exercise: repeat with \(S_5\).

\subsection{Conjugation in Alternating Groups}
Note that
\begin{align*}
	\ccl_{S_n}(\sigma) & = \{ \tau \sigma \tau^{-1} : \tau \in S_n \} \\
	\ccl_{A_n}(\sigma) & = \{ \tau \sigma \tau^{-1} : \tau \in A_n \}
\end{align*}
So clearly \(\ccl_{A_n}(\sigma) \subseteq \ccl_{S_n}(\sigma)\) since \(A_n \subseteq S_n\).
But elements that are conjugate in \(S_n\) may not be conjugate in \(A_n\), for example \((1\ 2\ 3)\) and \((1\ 3\ 2)\) in \(S_3\) and \(A_3\).
