\subsection{All Transformations in $O_2$}
\begin{theorem}
	Every element of $SO_2$ is of the form
	\[ \begin{pmatrix}
			\cos\theta & -\sin\theta \\
			\sin\theta & \cos\theta
		\end{pmatrix} \]
	for some $\theta \in [0, 2\pi)$. This is an anticlockwise rotation of $\mathbb R^2$ about the origin by angle $\theta$. Conversely, every such element lies in $SO_2$.
\end{theorem}
\begin{proof}
	Let
	\[ A = \begin{pmatrix}
			a & b \\ c & d
		\end{pmatrix} \in SO_2 \]
	We have $A^\transpose A = I$ and $\det A = 1$. So
	\[ A^\transpose = A^{-1} \implies \begin{pmatrix}
			a & c \\ b & d
		\end{pmatrix} = \frac{1}{1} \begin{pmatrix}
			d & -b \\ -c & a
		\end{pmatrix} \]
	So $a=d, b=-c$. Since $ad-bc=1$, $a^2+c^2=1$. Then we can write $a = \cos \theta$ and $c = \sin \theta$ for a unique $\theta \in [0, 2\pi)$. Conversely, the determinant of this matrix is 1, and is in $O_2$, so this element lies in $SO_2$.
\end{proof}
\begin{theorem}
	The elements of $O_2 \setminus SO_2$ are the reflections in lines through the origin.
\end{theorem}
\begin{proof}
	Let
	\[ A = \begin{pmatrix}
			a & b \\ c & d
		\end{pmatrix} \in O_2 \setminus SO_2 \]
	So $A^\transpose A = I$ and $\det A = -1$.
	\[ A^\transpose = A^{-1} \implies \begin{pmatrix}
			a & c \\ b & d
		\end{pmatrix} = \frac{1}{-1} \begin{pmatrix}
			d & -b \\ -c & a
		\end{pmatrix} \]
	So $a=-d, b=c$. Together with $ad-bc=-1$, we have $a^2 + c^2 = 1$. So let $a = \cos \theta$, $c = \sin \theta$ like before, so
	\[ A = \begin{pmatrix}
			\cos \theta & \sin \theta  \\
			\sin \theta & -\cos \theta
		\end{pmatrix} \]
	which can be shown to be a reflection using double angle formulas such that
	\[ A \begin{pmatrix}
			\sin \frac{\theta}{2} \\ \cos \frac{\theta}{2}
		\end{pmatrix} = -\begin{pmatrix}
			\sin \frac{\theta}{2} \\ \cos \frac{\theta}{2}
		\end{pmatrix};\quad A\begin{pmatrix}
			\cos \frac{\theta}{2} \\ \sin \frac{\theta}{2}
		\end{pmatrix} = \begin{pmatrix}
			\cos \frac{\theta}{2} \\ \sin \frac{\theta}{2}
		\end{pmatrix} \]
	So $A$ is a reflection in the plane orthogonal to the vector $\begin{pmatrix}
			\sin \frac{\theta}{2} \\ \cos \frac{\theta}{2}
		\end{pmatrix}$. Conversely, any reflection in a line through the origin has this form, so it will be in $O_2 \setminus SO_2$.
\end{proof}

\begin{corollary}
	Every element of $O_2$ is the composition of at most two reflections.
\end{corollary}
\begin{proof}
	Every element of $O_2 \setminus SO_2$ is a reflection, so this is trivial. If $A \in SO_2$, then we can write
	\[ A = \underbrace{A \begin{pmatrix}
				-1 & 0 \\ 0 & 1
			\end{pmatrix}}_{\det = -1} \underbrace{\begin{pmatrix}
				-1 & 0 \\ 0 & 1
			\end{pmatrix}}_{\det = -1} \]
	So we have expressed $A$ as the product of two reflections.
\end{proof}

\subsection{All Transformations in $O_3$}
\begin{theorem}
	If $A \in SO_3$, then there exists some unit vector $\vb v \in \mathbb R^3$ with $A\vb v = \vb v$, i.e. there exists an eigenvector with eigenvalue 1.
\end{theorem}
\begin{proof}
	It is sufficient to show that 1 is an eigenvalue of $A$, since this guarantees that there is some nonzero eigenvector for this eigenvalue which we can then normalise. This is equivalent to showing that $\det (A - I) = 0$.
	\begin{align*}
		\det(A - I) & = \det(A - AA^\transpose)       \\
		            & = \det(A)\det(I - A^\transpose) \\
		            & = \det(I - A^\transpose)        \\
		            & = \det((I - A)^\transpose)      \\
		            & = \det(I - A)                   \\
		            & = (-1)^3\det(A - I)
	\end{align*}
	So $2\det(A - I) = 0 \implies \det(A - I) = 0$.
\end{proof}
\begin{corollary}
	Every element $A \in SO_3$ is conjugate (in $SO_3$) to a matrix of the form
	\[ \begin{pmatrix}
			1 & 0           & 0            \\
			0 & \cos \theta & -\sin \theta \\
			0 & \sin \theta & \cos \theta
		\end{pmatrix} \]
\end{corollary}
\begin{proof}
	By the above theorem, there exists some unit vector $\vb v_1$ which is an eigenvector of eigenvalue 1. We can extend this vector to an orthonormal basis $\{ \vb v_1, \vb v_2, \vb v_3 \}$ of $\mathbb R^3$. Then, for $i=2,3$, we have
	\[ A\vb v_i \cdot \vb v_1 = A\vb v_i \cdot A\vb v_1 = \vb v_i \cdot \vb v_1 = 0 \]
	So $A\vb v_2, A\vb v_3$ lie in the subspace generated by $\vb v_2, \vb v_3$, i.e. $\vecspan \{ \vb v_2, \vb v_3 \} = \genset{\vb v_2, \vb v_3}$. So $A$ maps this subspace to itself, and we can thus consider the restriction of $A$ to this subspace. The matrix in this new basis will have form
	\[ \begin{pmatrix}
			1 & 0 & 0 \\
			0 & a & b \\
			0 & c & d
		\end{pmatrix} \]
	The smaller matrix in the bottom right will still have determinant 1, since we can expand the determinant here by the first row. So $A$ restricted to this subspace is an element of $SO_2$, so its matrix must be of the form
	\[ \begin{pmatrix}
			a & b \\ c & d
		\end{pmatrix} = \begin{pmatrix}
			\cos \theta & -\sin \theta \\
			\sin \theta & \cos \theta
		\end{pmatrix} \]
	So $A$ has the required form with respect to this new basis $\{ \vb v_1, \vb v_2, \vb v_3 \}$. The change of basis matrix $P$ lies in $O_3$ since $\{ bm v_1, \vb v_2, \vb v_3 \}$ is an orthonormal basis. If $P \notin SO_3$, then we can use the basis $\{ -\vb v_1, \vb v_2, \vb v_3 \}$ instead, which will invert the determinant of $P$. So in either case $P \in SO_3$.
\end{proof}
This tells us in particular that every element in $SO_3$ is a rotation about some axis, here $\vb v_1$.

\begin{corollary}
	Every element of $O_3$ is the composition of at most three reflections.
\end{corollary}
\begin{proof}
	\begin{itemize}
		\item If $A \in SO_3$, then $\exists P \in SO_3$ such that $PAP^{-1} = B$, where $B$ is of the form
		      \[ B = \begin{pmatrix}
				      1 & 0           & 0            \\
				      0 & \cos \theta & -\sin \theta \\
				      0 & \sin \theta & \cos \theta
			      \end{pmatrix} \]
		      Since this smaller matrix
		      \[ \begin{pmatrix}
				      \cos \theta & -\sin \theta \\
				      \sin \theta & \cos \theta
			      \end{pmatrix} \]
		      is a composition of at most two reflections, then $B$ is also a composition of at most two reflections, i.e. $B = B_1 B_2$. Since $A$ is a conjugate of $B$, it is also a composition of at most two reflections, as the conjugate of a reflection is a reflection, and $A = P^{-1}BP = (P^{-1}B_1P)(P^{-1}B_2P)$.
		\item If $A \in O_3 \setminus SO_3$, then $\det A = -1$ and we can construct
		      \[ A = \underbrace{A\begin{pmatrix}
					      -1 & 0 & 0 \\
					      0  & 1 & 0 \\
					      0  & 0 & 1
				      \end{pmatrix}}_{\det = 1}\underbrace{\begin{pmatrix}
					      -1 & 0 & 0 \\
					      0  & 1 & 0 \\
					      0  & 0 & 1
				      \end{pmatrix}}_{\det = -1} \]
		      So the left-hand product lies in $SO_3$, so it is a composition of at most two reflections. The final element is a reflection in the $y$--$z$ plane, so the combined product is a composition of at most three reflections.
	\end{itemize}
\end{proof}

\subsection{Symmetries of the Cube (revisited)}
We can think of symmetry groups of the Platonic solids as subgroups of $O_3$ by placing the solid at the origin. By question 11 on example sheet 4, we have that $O_3 \cong SO_3 \times C_2$, where $C_2$ is generated by the map $\vb v \mapsto -\vb v$. So if $\vb v\mapsto -\vb v$ is a symmetry of our platonic solid, then this group of symmetries will also split as the direct product of $G^+ \times C_2$ where $G^+$ is the group of rotations (proof as exercise).

So we have that the group of symmetries of the cube is $G^+ \times C_2 \cong S_4 \times C_2$ by the results from earlier.
