\subsection{Definition}
\begin{definition}
	A function \( c \colon V(G) \to \qty{1, \dots, k} \) is a \emph{(proper) \( k \)-colouring} of a graph if \( x \sim y \implies c(x) \neq c(y) \).
	The \emph{chromatic number} of \( G \), denoted \( \chi(G) \), is the minimum \( k \) such that there exists a \( k \)-colouring of \( G \).
\end{definition}
\begin{example}
	A path \( P_n \) has a 2-colouring.
	More generally, a graph is bipartite if and only if it has a 2-colouring.
	An even cycle has chromatic number 2, and an odd cycle has chromatic number 3.
	A tree has chromatic number 2.
	The complete graph on \( n \) vertices has chromatic number \( n \).
\end{example}
\begin{proposition}
	Let \( G \) be a graph.
	Then \( \chi(G) \leq \Delta(G) + 1 \).
\end{proposition}
\begin{proof}
	Let \( x_1, \dots, x_n \) be an ordering of the vertices of \( G \).
	We create a colouring of the vertices by induction.
	Suppose \( x_1, \dots, x_i \) have already been coloured, and we want to colour \( x_{i+1} \).
	Since \( x_{i+1} \) has at most \( \Delta(G) \) neighbours that have already been coloured, but we have \( \Delta(G) + 1 \) available colours, there is a free colour that does not match any previous neighbours.
	Choose the smallest available colour.
	By induction we can colour the entire graph.
\end{proof}
\begin{remark}
	This is sometimes known as a \emph{greedy colouring}.
	The greedy colouring may produce a colouring which is suboptimal for a given graph; consider the path \( P_4 \) on the vertex set \( \qty{1, 2, 3, 4} \) but with the ordering \( 1, 4, 2, 3 \): this gives a 3-colouring.
	The proposition above is sharp: the chromatic number of the complete graph is \( n \), and its maximum degree is \( n - 1 \).
\end{remark}

\subsection{Colouring planar graphs}
\begin{proposition}
	Let \( G \) be planar.
	Then \( \delta(G) \leq 5 \).
\end{proposition}
\begin{proof}
	The average degree of \( G \), given by \( n^{-1} \sum_{v \in V(G)} \deg v \), is exactly \( 2 n^{-1} e(G) \).
	Since \( e(G) \leq 3n - 6 \), the average degree at most \( 6 - \frac{12}{n} < 6 \), so \( \delta(G) \leq 5 \).
\end{proof}
\begin{proposition}[six-colour theorem]
	Let \( G \) be planar.
	Then \( G \) admits a 6-colouring.
\end{proposition}
\begin{proof}
	Apply induction on \( \abs{G} \).
	If \( \abs{G} \leq 6 \), there admits a trivial 6-colouring.
	Let \( G \) be planar, and let \( x \in V(G) \) have degree at most 5.
	By the inductive hypothesis, \( G - x \) admits a 6-colouring.
	Since \( x \) has at most five neighbours, there is a free colour to use for \( x \).
\end{proof}
\begin{theorem}[five-colour theorem]
	Let \( G \) be planar.
	Then \( G \) admits a 5-colouring.
\end{theorem}
\begin{proof}
	We apply induction on \( \abs{G} \).
	Clearly the theorem holds for \( \abs{G} \leq 5 \).
	Suppose \( \abs{G} > 5 \).
	Let \( x \in V(G) \) be a vertex with degree at most five.
	Applying induction, there exists a 5-colouring of \( G - x \).
	If the degree is four or lower, we can use the free colour to colour \( x \), so suppose \( x \) has degree five.
	Let \( N(x) = \qty{x_1, x_2, x_3, x_4, x_5} \) arranged cyclically in the plane, and let the colour of \( x_i \) be \( i \).
	Then without loss of generality, all the \( x_i \) must have different colours, since otherwise, we are done.

	Suppose there exists no path from \( x_1 \) to \( x_3 \) in \( G - x \) only along vertices coloured \( 1 \) and \( 3 \).
	In this case, let \( C \) be the component of \( G \) of vertices coloured \( 1 \) or \( 3 \) that contains \( x_1 \).
	This is the connected component of the subgraph of \( G - x \) induced by the vertices coloured \( 1 \) and \( 3 \) that contains \( x_1 \).
	By assumption, \( x_3 \) is not in this component.
	Now, swap the colours \( 1 \) and \( 3 \) on \( C \); this yields another 5-colouring of \( G - x \).
	We can then extend this 5-colouring to \( x \) by colouring \( x \) with \( 1 \).

	Now, suppose there exists no path from \( x_2 \) to \( x_4 \) in \( G - x \) along vertices coloured \( 2 \) and \( 4 \).
	If so, we are done as above.

	Suppose that there exists an \( x_1 \)--\( x_3 \) path using only colours \( 1 \) and \( 3 \), and an \( x_2 \)--\( x_4 \) path using only colours \( 2 \) and \( 4 \).
	Then since both paths lie in the plane and the vertices are arranged cyclically as above, they must cross.
	The intersection vertex is coloured either \( 1 \) or \( 3 \), and also either \( 2 \) or \( 4 \).
	This is a contradiction.
\end{proof}
\begin{remark}
	Any planar graph admits a 4-colouring; this result is known as the \emph{four-colour theorem}.
	The above method does not work, because when swapping the colours of a component, there is not a free colour to use for the newly added vertex.
	The four-colour theorem was eventually proven using a computer-aided search after reducing the problem to thousands of specific local configurations.
	The four-colour theorem is sharp; \( K_4 \) is planar.
\end{remark}

\subsection{Colouring non-planar graphs}
\begin{proposition}
	Let \( G \) be a connected graph, and \( \delta(G) < \Delta(G) \).
	Then \( \chi(G) \leq \Delta(G) \).
\end{proposition}
\begin{proof}
	Order the vertices in \( G \) into \( x_1, \dots, x_n \) such that \( \deg x_n < \Delta(G) \), and \( x_{n-1} \) is adjacent to \( x_n \), and also \( x_{n-2} \) is adjacent to one of \( x_n \) and \( x_{n-1} \) and so on.
	This is always possible since \( G \) is connected.
	This ordering has the property that all vertices have less than \( \Delta(G) \) edges facing forward.
	So the greedy colouring gives a \( \Delta(G) \)-colouring.
\end{proof}
\begin{theorem}[Brooks]
	Let \( G \) be a connected graph.
	If \( G \) is not an odd cycle or complete graph, \( \chi(G) \leq \Delta(G) \).
\end{theorem}
\begin{remark}
	We have shown above that \( \chi(G) \leq \Delta(G) + 1 \).
	This theorem then says that \( \chi(G) = \Delta(G) + 1 \) if and only if \( G \) is an odd cycle or a complete graph.
\end{remark}
\begin{proof}
	We apply induction on \( \abs{G} \).
	We can check that the theorem holds for \( \abs{G} \leq 3 \).
	Note that we may assume that \( \Delta(G) \geq 3 \); otherwise, the graph is bipartite or an odd cycle.

	We will show first that if \( G \) is 3-connected, the theorem holds.
	We give an ordering of \( V(G) \).
	Let \( x_n \) be a vertex of degree \( \Delta(G) \), and let \( x_1, x_2 \in N(x) \) be non-adjacent vertices.
	This is possible; indeed, suppose we could not find such vertices.
	Then \( \qty{x} \cup N(x) \) is a complete graph, so \( G = K_{\Delta(G) + 1} \) by connectedness, contradicting our assumption.
	Now, consider \( G - \qty{x_1, x_2} \).
	Since \( G \) is 3-connected, \( G - \qty{x_1,x_2} \) is connected.
	We can order the vertices in the same way as above, choosing \( x_{n-1} \sim x_n \) and \( x_{n-2} \) a neighbour of \( x_{n-1} \) or \( x_n \), and so on.
	Then the greedy algorithm produces the required colouring.

	Now, we show that if \( \kappa(G) = 1 \), the theorem holds.
	In this case, we have a separator of size one, so let \( \qty{x} \) be such a separator (we call \( x \) a \emph{cut vertex}).
	Let \( C_1, \dots, C_n \) be the connected components of \( G - x \).
	By induction, we can colour \( C_i \cup \qty{x} \) for each \( i \); they cannot be complete, by counting the number of edges of \( x \) in this graph.
	We can then permute the colours in each such colouring to make \( x \) the same colour.
	Then we can combine each colouring to produce a colouring of the entire graph.

	Finally, we will consider the case when \( \kappa(G) = 2 \).
	Let \( S = \qty{x,y} \) be a separator for \( G \).
	Let \( C_1, \dots, C_k \) be the components of \( G - S \).
	Define the graphs \( G_i = G[C_i \cup S] + xy \) for \( i = 1, \dots, k \).

	Suppose \( \delta(G_i) < \Delta(G) \) for all \( i \).
	In this case, the \( G_i \) can be coloured by induction as they are not complete graphs.
	Note that \( x, y \) get different colours since we have added the edge \( xy \).
	Therefore, we can permute the colours, such that the colouring agrees on \( x,y \) for all \( G_i \).
	These colourings can be combined into a \( \Delta(G) \)-colouring of \( G \).

	Now suppose without loss of generality that \( \delta(G_1) = \Delta(G) \).
	In this case, \( k = 2 \), and \( \abs{N(x) \cap C_1} = \Delta(G) - 1 = \abs{N(x) \cap C_1} \), and \( \abs{N(x) \cap C_2} = 1 = \abs{N(y) \cap C_2} \).
	Let \( x', y' \) be the neighbours of \( x, y \) in \( C_2 \).
	Now, note that \( \widetilde S = \qty{x,y'} \) is a separator, and now \( \delta(G_i) < \Delta(G) \) for all connected components, and we can use the proof from above.
\end{proof}

\subsection{Chromatic polynomial}
\begin{definition}
	Let \( G \) be a graph.
	The \emph{chromatic polynomial} of \( G \) is \( P_G \colon \mathbb Z_{\geq 0} \to \mathbb Z_{\geq 0} \) where \( P_G(t) \) is the number of \( t \)-colourings of \( G \).
\end{definition}
\begin{remark}
	The minimum \( t \) for which \( P_G(t) > 0 \) the chromatic number.
\end{remark}
\begin{example}
	The chromatic polynomial on the empty graph on \( n \) vertices is given by \( P_G(t) = t^n \).

	The chromatic polynomial on the complete graph on \( n \) vertices is \( P_G(t) = t(t-1)\dots(t-(n-1)) = \binom t n \).

	For a path on \( n \) vertices, \( P_G(t) = t(t-1)^{n-1} \).
	For any tree, colouring each leaf, removing it, then colouring the remainder inductively, \( P_G(t) = t(t-1)^{\abs{G} - 1} \).
\end{example}
\begin{definition}
	Let \( G \) be a graph, and \( e = xy \in E(G) \).
	The \emph{contraction} of \( G \) along \( e \), denoted \( G / e \), is the graph with vertices \( V(G) \setminus \qty{x, y} \cup \qty{a} \) for a new variable \( a \), and edges \( E(G[V\setminus\qty{x,y}]) \cup \qty{az \mid x \sim z} \cup \qty{az \mid y \sim z} \).
\end{definition}
\begin{proposition}
	Let \( G \) be a graph and \( e \in E(G) \).
	Then \( P_G(t) = P_{G - e}(t) - P_{G/e}(t) \).
\end{proposition}
\begin{proof}
	Let \( e = xy \).
	A \( t \)-colouring of \( G - e \) where \( x, y \) are assigned different colours corresponds to a \( t \)-colouring of \( G \), by simply adding the edge back.
	A \( t \)-colouring of \( G - e \) where \( x, y \) are assigned the same colour corresponds to a \( t \)-colouring of \( G / e \), by contracting the edge.
\end{proof}
\begin{remark}
	The above proposition is known as a `cut-fuse' relation.
\end{remark}
\begin{proposition}
	Let \( G \) be a graph.
	Then \( P_G(t) \) is indeed a polynomial with degree \( \abs{G} \).
\end{proposition}
\begin{proof}
	We apply induction on \( e(G) \).
	If there are no edges in the graph, the graph is empty, and has chromatic polynomial \( P_G(t) = t^{\abs{G}} \).
	Otherwise, let \( e \in E(G) \).
	By induction, \( P_{G - e}(t) \) is a polynomial of degree \( \abs{G - e} = \abs{G} \), and \( P_{G / e}(t) \) is a polynomial of degree \( \abs{G / e} = \abs{G} - 1 \).
	Hence \( P_G(t) = P_{G - e}(t) - P_{G / e}(t) \) is indeed a polynomial of the required degree.
\end{proof}
\begin{proposition}
	Let \( G \) be a graph with \( n \) vertices and \( m \) edges.
	Then \( P_G(t) = t^n - mt^{n-1} + p(t) \) where \( p \) is a polynomial of degree at most \( n - 2 \).
\end{proposition}
\begin{proof}
	We apply induction on \( e(G) \).
	If there are no edges, we have the empty graph, which has the required form.
	Otherwise, let \( e \in E(G) \).
	Then
	\[ P_G(t) = P_{G - e}(t) - P_{G/e}(t) = \qty(t^n - (m - 1)t^{n-1} + \dots) + \qty(t^{n-1} + \dots) = t^n - mt^{n-1} + \dots \]
	as required.
\end{proof}
\begin{remark}
	Other coefficients of the chromatic polynomial contain other information about the graph.
	For example, the \( t^{n-2} \) coefficient is exactly \( \binom n 2 - \text{ number of triangles in } G \).

	If \( G \) is planar, \( P_G(2+\frac{1+\sqrt{5}}{2}) \neq 0 \).

	A result due to June Huh is that the coefficients \( c_0, \dots, c_n \) of \( P_G \) are \emph{log-concave}, so \( c_i^2 > c_{i-1}c_{i+1} \).
\end{remark}
