\subsection{Definition}
\begin{definition}
	A tensor is isotropic if it is invariant under changes with respect to the choice of Cartesian coordinate axes.
	\[
		T_{ij\dots k}' = R_{ip} R_{jq} \dots R_{kr} T_{pq\dots r} = T_{ij\dots k}
	\]
	for any choice of rotation \(R\).
\end{definition}
\noindent Note that by definition, every scalar is isotropic.
The Kronecker and Levi-Civita tensors are also isotropic, as we saw above.

\subsection{Classifying isotropic tensors in \(\mathbb R^3\)}
\begin{proposition}
	The isotropic tensors on \(\mathbb R^3\), ordered by rank, are exactly (up to the multiplication of a multiplicative scalar)
	\begin{enumerate}[{Rank} 1:]
		\setcounter{enumi}{-1}
		\item all tensors
		\item no non-zero tensors
		\item the Kronecker \(\delta\)
		\item the Levi-Civita \(\varepsilon\)
		\item \(\alpha \delta_{ij} \delta_{k\ell} + \beta \delta_{ik} \delta_{j\ell} + \gamma \delta_{i\ell} \delta_{jk}\) where \(\alpha\), \(\beta\), \(\gamma\) are scalars
	\end{enumerate}
	and for ranks higher than 4, they are a linear combination of products of \(\delta\) and \(\varepsilon\) terms, for instance \(\delta_{ij}\varepsilon_{k\ell m}\).
\end{proposition}
\begin{proof}
	This is a non-rigorous sketch proof.
	\begin{enumerate}[{Rank} 1:]
		\setcounter{enumi}{-1}
		\item By definition, such tensors do not transform components under a change of basis.
		\item Let \(v_i\) be the components of an isotropic vector of rank 1.
		      Then, for any \(R\), we must have
		      \[
			      v_i = R_{ij} v_j
		      \]
		      Let \(R\) be a rotation by \(\pi\) about the \(z\) axis, so
		      \[
			      R = \begin{pmatrix}
				      -1 & 0  & 0 \\
				      0  & -1 & 0 \\
				      0  & 0  & 1
			      \end{pmatrix}
		      \]
		      Hence,
		      \[
			      v_1 = -v_1;\quad v_2 = -v_2;\quad v_3 = v_3
		      \]
		      Hence, \(v_1 = 0, v_2 = 0\).
		      Alternatively, let
		      \[
			      R = \begin{pmatrix}
				      1 & 0  & 0  \\
				      0 & -1 & 0  \\
				      0 & 0  & -1
			      \end{pmatrix}
		      \]
		      Then clearly \(v_3 = -v_3 = 0\).
		      Hence the only tensor with this property is the zero tensor.
		\item If \(T_{ij}\) are the components of an isotropic tensor of rank 2, then for all choices of \(R\), we have
		      \[
			      T_{ij} = R_{ip} R_{jq} T_{pq}
		      \]
		      Let \(R\) be a rotation by \(\frac{\pi}{2}\) about each axis, so for example in the \(z\) direction,
		      \[
			      R = \begin{pmatrix}
				      0  & 1 & 0 \\
				      -1 & 0 & 0 \\
				      0  & 0 & 1
			      \end{pmatrix}
		      \]
		      So \(T_{13} = R_{1p} R_{3q} T_{pq} = R_{12} R_{33} T_{23} = T_{23}\).
		      Analogously we find, \(T_{23} = -T_{13}\).
		      Hence, \(T_{13} = T_{23} = 0\).
		      Further, \(T_{11} = R_{1p} R_{1q} T_{pq} = R_{12} R_{12} T_{22} = T_{22}\).
		      So by symmetry,
		      \[
			      T_{11} = T_{22} = T_{33};\quad T_{13} = T_{23} = T_{12} = T_{31} = T_{32} = T_{21} = 0
		      \]
		      which is exactly the \(\delta\) tensor, up to a scale factor.
		\item For rank 3 tensors, we can use the same idea, but with more indices.
	\end{enumerate}
\end{proof}

\subsection{Integrals with isotropic tensors}
Consider an integral of the form
\[
	T_{ij\dots k} = \int_{\abs{\vb x} < R} f(r) x_i x_j \dots x_k \dd{V}
\]
where \(x_k x_k = r^2\), and \(\dd{V(x)} = \dd{x_1} \dd{x_2} \dd{x_3}\).
Note that \(f(r)\) and \(\{ \vb x \colon \abs{\vb x} < R \}\) are invariant under rotation.
Since \(\abs{J}\) under a rotation is 1, we have
\begin{align*}
	T_{ij\dots k}' & = \int_{\abs{\vb x} < R} f(r) x_i' x_j' \dots x_k' \dd{x_1'} \dd{x_2'} \dd{x_3'}                \\
	               & = \int_{\abs{\vb x} < R} f(r) R_{ip} x_p R_{jq} x_q \dots R_{kr} x_r \dd{x_1} \dd{x_2} \dd{x_3}
\end{align*}
We will now make the substitution
\[
	y_i = R_{ij} x_j;\quad \dd{V} = \dd{y_1} \dd{y_2} \dd{y_3}
\]
Hence,
\begin{align*}
	T_{ij\dots k}' & = \int_{\abs{\vb x} < R} f(r) y_i y_j \dots y_k \dd{V(\vb y)} \\
	               & = \int_{\abs{\vb x} < R} f(r) x_i x_j \dots x_k \dd{V(\vb x)} \\
	               & = T_{ij\dots k}
\end{align*}
Hence such an integral always yields an isotropic tensor.
If we take \(R \to \infty\), this corresponds to an integral over \(\mathbb R^3\).
As an example, consider
\[
	T_{ij} = \int_{\mathbb R^3} e^{-r^5}x_i x_j \dd{V}
\]
Then \(T_{ij}\) is isotropic, hence \(T_{ij} = \alpha \delta_{ij}\).
Contracting on \((i, j)\) to find \(\alpha\), we get
\begin{align*}
	\alpha \delta_{ii} & = 3\alpha                                      \\
	                   & = \int_{\mathbb R^3} e^{-r^5} r^2 \dd{V}       \\
	                   & = 4\pi \int_{0}^\infty e^{-r^5} r^2 r^2 \dd{r} \\
	                   & = 4\pi \int_{0}^\infty e^{-r^5} r^4 \dd{r}     \\
	                   & = \frac{4}{5}\pi
\end{align*}
Hence,
\[
	T_{ij} = \frac{4}{15}\pi\delta_{ij}
\]
As another example, consider the inertia tensor \(I_{ij}\) of a ball of radius \(R\), uniform density \(\rho_0\), and mass \(M = \frac{4\pi}{3}R^3 \rho_0\).
Recall that
\[
	I_{ij} = \int_{\abs{\vb x} < R} \rho_0 (x_k x_k \delta_{ij} - x_i x_j) \dd{V}
\]
Both terms give an isotropic result, so the sum \(I_{ij}\) is isotropic.
Contracting on \((i, j)\), we have
\begin{align*}
	\alpha \delta_{ii} & = 3\alpha                                                                  \\
	                   & = \int_{\abs{\vb x} < R} \rho_0 (r^2 \delta_{ii} - x_i x_i) \dd{V}         \\
	                   & = \int_{\abs{\vb x} < R} \rho_0 (3r^2 - r^2) \dd{V}                        \\
	                   & = \int_{\abs{\vb x} < R} \rho_0 2r^2 \dd{V}                                \\
	                   & = 4\pi \int_0^R \rho_0 2r^4 \dd{r}                                         \\
	                   & = \frac{4\pi}{3}\rho_0 R^3 \qty(\frac{3}{R^3} \cdot 2 \cdot \frac{R^5}{5}) \\
	                   & = \frac{6MR^2}{5}
\end{align*}
Hence,
\[
	I_{ij} = \frac{2MR^2}{5}\delta_{ij}
\]
